---
title: "Data Cleaning & Alignment (EI/Sienna ‚ÜĒ EIA)"
author: "Sara Abril-Guevara"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: true
    toc_float: true
    df_print: paged
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) 
library(readr) 
library(ggplot2) 
library(tidyr) 
library(knitr)
library(readxl)
library(stringr)
library(dplyr)
library(readr)
library(stringdist)
library(purrr)
```

# Overview

This document summarizes the data cleaning and validation process used to align generator fleets from EI/Sienna and EIA datasets for selected states (MN, IL, MD, VA, WV).

The goal is to ensure consistency in installed capacity before running simulation models.


## Methodology

- DONE: Extract EI/Sienna generators and bus attributes using Kestrel

- DONE: Aggregate generators by geographic coordinates

- DONE: Standardize technology categories across datasets

- WIP: Compare installed capacity by state and technology

- WIP: Flag differences for further spatial matching and manual review

```{r cars, message = FALSE}
# EI datasets 
VA_ei_gens = read_csv("~/Downloads/virginia_resources.csv") 
MD_ei_gens = read_csv("~/Downloads/maryland_resources.csv")
WV_ei_gens = read_csv("~/Downloads/west_virginia_resources.csv")

#EIA
eia_all <- read_excel("3_1_Generator_Y2024.xlsx", sheet = "Operable", skip = 1, col_types = "text")

eia_all <- eia_all %>%
  mutate(
    `Nameplate Capacity (MW)` = as.numeric(`Nameplate Capacity (MW)`),
    `Summer Capacity (MW)` = as.numeric(`Summer Capacity (MW)`),
    `Winter Capacity (MW)` = as.numeric(`Winter Capacity (MW)`),
    `Minimum Load (MW)` = as.numeric(`Minimum Load (MW)`),
    `Nameplate Power Factor` = as.numeric(`Nameplate Power Factor`)
  ) %>%
  rename(nameplate_capacity = `Nameplate Capacity (MW)`)
eia_va <- eia_all %>% filter(State == "VA")
eia_md <- eia_all %>% filter(State == "MD")
eia_wv <- eia_all %>% filter(State == "WV")


# # EIA datasets 
# # Renewable 
# eia_va_renew = read_csv("~/Documents/GitHub/va_updates/renewable_VA_EIA.csv") 
# # Hydro 
# eia_va_hydro = read_csv("~/Documents/GitHub/va_updates/va_hydro_EIA.csv") 
# # Thermal 
# eia_va_thermal = read_excel("~/Documents/GitHub/va_updates/Thermal_EIA.xlsx")
# 
# # Combine all VA EIA datasets 
# eia_va = bind_rows( eia_va_renew, eia_va_hydro, eia_va_thermal )
# 
# # Renewable 
# eia_md_renew = read_csv("~/Documents/GitHub/md_update/Renewable_EIA.csv") 
# # Hydro 
# eia_md_hydro = read_csv("~/Documents/GitHub/md_update/Hydro_EIA.csv")
# # Thermal 
# eia_md_thermal = read_csv("~/Documents/GitHub/md_update/Thermal_EIA.csv")
# eia_md_renew = eia_md_renew %>%
#   mutate(`generator ID` = as.character(`generator ID`))
# 
# eia_md_hydro = eia_md_hydro %>%
#   mutate(`generator ID` = as.character(`generator ID`))
# 
# eia_md_thermal = eia_md_thermal %>%
#   mutate(`generator ID` = as.character(`generator ID`))
# 
# eia_md = bind_rows( eia_md_renew, eia_md_hydro, eia_md_thermal )
# #eia_wv = read_csv("renewable_WV_EIA.csv")
```

```{r}
# clean_eia = function(df, state_name) {
#   df %>%
#     mutate(
#       # Fill technology using mode when missing
#       technology = case_when(
#         !is.na(technology) ~ technology,
#         mode == "Storage" ~ "Pumped Storage Hydroelectric",
#         mode == "RoR" ~ "Run-of-River Hydroelectric",
#         TRUE ~ "Unknown"
#       ),
#       
#       # Add state
#       state = state_name
#     )
# }
# eia_va = clean_eia(eia_va, "VA")
# eia_md = clean_eia(eia_md, "MD")
# 
# clean_eia_final = function(df) {
#   df %>%
#     select(
#       plant,
#       `generator ID`,
#       technology,
#       lat,
#       lon,
#       nameplate_capacity,
#       ba,
#       state
#     ) %>%
#     filter(!is.na(lat) & !is.na(lon))  # remove bad rows
# }
# 
# eia_va = clean_eia_final(eia_va)
# eia_md = clean_eia_final(eia_md)
```

```{r}
# Function to map EIA generators to high-level tech groups
map_tech_group <- function(df) {
  df %>%
    mutate(
      tech_group = case_when(
        # Nuclear
        `Prime Mover` == "NU" | Technology == "Nuclear" ~ "Nuclear",
        
        # Pumped Storage
        `Prime Mover` == "PS" | str_detect(Technology, "Pumped Storage") ~ "Pumped Storage",
        
        # Hydro (excluding Pumped Storage)
        `Prime Mover` == "HY" | str_detect(Technology, "Hydroelectric") & !str_detect(Technology, "Pumped Storage") ~ "Hydro",
        
        # Wind
        `Prime Mover` == "WT" | str_detect(Technology, "Wind") ~ "Wind",
        
        # Solar
        `Prime Mover` %in% c("PV", "CE") | str_detect(Technology, "Solar") ~ "Solar",
        
        # Thermal / Fossil
        `Prime Mover` %in% c("ST", "CT", "CA", "GT", "IC", "CS", "FC", "CP") |
          Technology %in% c(
            "Conventional Steam Coal", 
            "Natural Gas Fired Combined Cycle", 
            "Natural Gas Steam Turbine",
            "Coal Integrated Gasification Combined Cycle",
            "Natural Gas Fired Combustion Turbine",
            "Natural Gas Internal Combustion Engine"
          ) ~ "Thermal",
        
        # Batteries / Storage / Flywheels
        `Prime Mover` %in% c("BA", "BT", "FW") | str_detect(Technology, "Battery|Flywheels|Storage") ~ "Batteries",
        
        # Biomass / Waste / Geothermal / Other
        str_detect(Technology, "Biomass|Waste|Geothermal|Other") ~ "Other",
        
        # Fallback
        TRUE ~ "Other"
      )
    )
}
# Clean and map tech_group for all three states
eia_va <- eia_va %>% map_tech_group()
eia_md <- eia_md %>% map_tech_group()
eia_wv <- eia_wv %>% map_tech_group()
```




## Technology Mapping

```{r tech-mapping}
# map_technology = function(df, source = "EI") {
# 
#   if (source == "EI") {
#     df %>%
#       mutate(tech_group = case_when(
#         fuel == "NUCLEAR" ~ "Nuclear",
#         generator_type %in% c("Solar") ~ "Solar",
#         generator_type %in% c("Wind") ~ "Wind",
#         generator_type %in% c("Hydro") ~ "Hydro",
#         generator_type %in% c("ThermalStandard", "ThermalMultiStart") ~ "Thermal",
#         generator_type %in% c("Storage") ~ "Storage",
#         TRUE ~ "Other"
#       ))
# 
#   } else {
#     df %>%
#       mutate(tech_group = case_when(
#         grepl("Solar", technology) ~ "Solar",
#         grepl("Wind", technology) ~ "Wind",
#         grepl("Hydro", technology) ~ "Hydro",
#         grepl("Pumped Storage", technology) ~ "Storage",
#         grepl("Battery|Storage", technology) ~ "Storage",
#         grepl("Nuclear", technology) ~ "Nuclear",
#         grepl("Gas|Coal|Petroleum|Steam|Combustion", technology) ~ "Thermal",
#         grepl("Biomass|Waste|Landfill", technology) ~ "Other",
#         TRUE ~ "Other"
#       ))
#   }
# }
# 
# # Summarize function including units for both EI and EIA
# summarize_capacity_units = function(ei_df, eia_df, state_name) {
# 
#   ei_clean  = map_technology(ei_df, "EI")
#   eia_clean = map_technology(eia_df, "EIA")
# 
#   # EI summary
#   ei_summary = ei_clean %>%
#     group_by(tech_group) %>%
#     summarise(
#       EI_total_units = n(),
#       EI_available_units = sum(available == TRUE, na.rm = TRUE),
#       EI_total_capacity = sum(rating, na.rm = TRUE),
#       EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
#       .groups = "drop"
#     )
# 
#   # EIA summary
#   eia_summary = eia_clean %>%
#     group_by(tech_group) %>%
#     summarise(
#       EIA_units = n(),
#       EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
#       .groups = "drop"
#     )
# 
#   # Join and calculate differences
#   full_join(ei_summary, eia_summary, by = "tech_group") %>%
#     mutate(
#       state = state_name,
#       diff_available_capacity = EI_available_capacity - EIA_capacity,
#       pct_diff_available_capacity = ifelse(EIA_capacity > 0,
#                                            (diff_available_capacity / EIA_capacity) * 100,
#                                            NA),
#       diff_total_capacity = EI_total_capacity - EIA_capacity,
#       pct_diff_total_capacity = ifelse(EIA_capacity > 0,
#                                        (diff_total_capacity / EIA_capacity) * 100,
#                                        NA)
#     ) %>%
#     select(state, tech_group, EI_total_units, EI_available_units, EIA_units,
#            EI_total_capacity, EI_available_capacity, EIA_capacity,
#            diff_total_capacity, pct_diff_total_capacity,
#            diff_available_capacity, pct_diff_available_capacity)
# }

```

```{r}
map_technology = function(df, source = "EI") {
  
  if (source == "EI") {
    
    df %>%
      mutate(
        tech_group = case_when(
          fuel == "NUCLEAR" ~ "Nuclear",
          generator_type %in% c("Solar") ~ "Solar",
          generator_type %in% c("Wind") ~ "Wind",
          generator_type %in% c("ThermalStandard", "ThermalMultiStart") ~ "Thermal",
          generator_type %in% c("Storage") ~ "Storage",
          prime_mover_type == "PS" ~ "Pumped Storage",
          generator_type %in% c("Hydro") ~ "Hydro",
          TRUE ~ "Other"
        )
      )
    
  } else {
    
    df %>%
      mutate(
        tech_group = case_when(
          
          # --- Explicit first (avoid misclassification) ---
          Technology == "Nuclear" ~ "Nuclear",
          
          Technology == "Hydroelectric Pumped Storage" | `Prime Mover` == "PS" ~ "Pumped Storage",
          
          Technology == "Conventional Hydroelectric" ~ "Hydro",
          
          Technology == "Batteries" ~ "Batteries",
          
          Technology %in% c("Flywheels", "Natural Gas with Compressed Air Storage") ~ "Storage",
          
          # --- Pattern-based (your original strength) ---
          grepl("Solar", Technology) ~ "Solar",
          grepl("Wind", Technology) ~ "Wind",
          
          grepl("Hydro", Technology) & !grepl("Pumped Storage", Technology) ~ "Hydro",
          
          grepl("Gas|Coal|Petroleum|Steam|Combustion", Technology) ~ "Thermal",
          
          grepl("Biomass|Waste|Landfill|Geothermal", Technology) ~ "Other",
          
          TRUE ~ "Other"
        )
      )
  }
}
# Summarize function including units for both EI and EIA
summarize_capacity_units = function(ei_df, eia_df, state_name) {

  ei_clean  = map_technology(ei_df, "EI")
  eia_clean = map_technology(eia_df, "EIA")

  # EI summary
  ei_summary = ei_clean %>%
    group_by(tech_group) %>%
    summarise(
      EI_total_units = n(),
      EI_available_units = sum(available == TRUE, na.rm = TRUE),
      EI_total_capacity = sum(rating, na.rm = TRUE),
      EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      .groups = "drop"
    )

  # EIA summary
  eia_summary = eia_clean %>%
    group_by(tech_group) %>%
    summarise(
      EIA_units = n(),
      EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
      .groups = "drop"
    )

  # Join summaries and calculate differences
  summary_df = full_join(ei_summary, eia_summary, by = "tech_group") %>%
    mutate(
      state = state_name,
      diff_available_capacity = EI_available_capacity - EIA_capacity,
      pct_diff_available_capacity = ifelse(EIA_capacity > 0,
                                           (diff_available_capacity / EIA_capacity) * 100,
                                           NA),
      diff_total_capacity = EI_total_capacity - EIA_capacity,
      pct_diff_total_capacity = ifelse(EIA_capacity > 0,
                                       (diff_total_capacity / EIA_capacity) * 100,
                                       NA)
    ) %>%
    select(state, tech_group, EI_total_units, EI_available_units, EIA_units,
           EI_total_capacity, EI_available_capacity, EIA_capacity,
           diff_total_capacity, pct_diff_total_capacity,
           diff_available_capacity, pct_diff_available_capacity)

  # Optional: reorder tech groups
  tech_order = c("Nuclear", "Thermal", "Hydro", "Pumped Storage", "Storage", "Solar", "Wind", "Other")
  summary_df$tech_group = factor(summary_df$tech_group, levels = tech_order)
  summary_df = summary_df %>% arrange(tech_group)

  return(summary_df)
}
```



### VA 
```{r}
## Run for VA
va_summary_units = summarize_capacity_units(VA_ei_gens, eia_va, "VA")
va_summary_units
```

### MD

```{r}
md_summary_units = summarize_capacity_units(MD_ei_gens, eia_md, "MD")
md_summary_units
```

### WV

```{r}
wv_summary_units = summarize_capacity_units(WV_ei_gens, eia_wv, "WV")
wv_summary_units
```

## Capacity Summary Comparison

```{r summary-function}
summarize_capacity_by_state = function(ei_df, eia_df, state_name) {
  
  # ---- EI: split available vs unavailable ----
  ei_clean = map_technology(ei_df, "EI")
  
  ei_available = ei_clean %>%
    filter(available == TRUE)
  
  ei_unavailable = ei_clean %>%
    filter(available == FALSE)
  
  # ---- Aggregate EI ----
  ei_summary_available = ei_available %>%
    group_by(tech_group) %>%
    summarise(EI_available_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  ei_summary_total = ei_clean %>%
    group_by(tech_group) %>%
    summarise(EI_total_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  # ---- EIA ----
  eia_clean = map_technology(eia_df, "EIA")
  
  eia_summary = eia_clean %>%
    group_by(tech_group) %>%
    summarise(EIA_capacity = sum(nameplate_capacity, na.rm = TRUE), .groups = "drop")
  
  # ---- Merge ----
  summary = full_join(ei_summary_available, eia_summary, by = "tech_group") %>%
    full_join(ei_summary_total, by = "tech_group") %>%
    mutate(
      state = state_name,
      EI_available_capacity = replace_na(EI_available_capacity, 0),
      EI_total_capacity = replace_na(EI_total_capacity, 0),
      EIA_capacity = replace_na(EIA_capacity, 0),
      
      # Main comparison (correct one)
      diff_available = EI_available_capacity - EIA_capacity,
      pct_diff_available = ifelse(EIA_capacity > 0,
                                 diff_available / EIA_capacity * 100, NA),
      
      # Diagnostic comparison
      diff_total = EI_total_capacity - EIA_capacity,
      pct_diff_total = ifelse(EIA_capacity > 0,
                             diff_total / EIA_capacity * 100, NA)
    )
  
  return(summary)
}
```

```{r}
va_summary = summarize_capacity_by_state(VA_ei_gens, eia_va, "VA")
md_summary = summarize_capacity_by_state(MD_ei_gens, eia_md, "MD")
wv_summary = summarize_capacity_by_state(WV_ei_gens, eia_wv, "WV")
all_states_summary = bind_rows(va_summary, md_summary, wv_summary)
```

```{r}
plot_thermal_discrepancy <- function(ei_df, eia_df, state_name, available_only = TRUE) {
  
  # -------------------------
  # EI Thermal (EXCLUDE NUCLEAR)
  # -------------------------
  ei_thermal <- ei_df %>%
    filter(prime_mover_type %in% c("ST", "CT", "CC", "OT")) %>%
    filter(fuel != "NUCLEAR") %>%   # <-- key fix
    { if (available_only) filter(., available == TRUE) else . } %>%
    group_by(fuel) %>%
    summarise(EI_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  # -------------------------
  # EIA Thermal
  # -------------------------
  eia_thermal <- eia_df %>%
    filter(Technology %in% c(
      "Conventional Steam Coal",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Petroleum Liquids",
      "Petroleum Coke",
      "Other Gases"
    )) %>%
    mutate(fuel = case_when(
      Technology == "Conventional Steam Coal" ~ "COAL",
      
      Technology %in% c(
        "Natural Gas Fired Combined Cycle",
        "Natural Gas Steam Turbine",
        "Natural Gas Fired Combustion Turbine",
        "Natural Gas Internal Combustion Engine",
        "Other Gases"
      ) ~ "NATURAL_GAS",
      
      Technology %in% c(
        "Petroleum Liquids",
        "Petroleum Coke"
      ) ~ "DISTILLATE_FUEL_OIL",
      
      TRUE ~ "OTHER"
    )) %>%
    group_by(fuel) %>%
    summarise(EIA_capacity = sum(nameplate_capacity, na.rm = TRUE), .groups = "drop")
  
  # -------------------------
  # Combine
  # -------------------------
  thermal_compare <- full_join(ei_thermal, eia_thermal, by = "fuel") %>%
    replace_na(list(EI_capacity = 0, EIA_capacity = 0)) %>%
    pivot_longer(cols = c(EI_capacity, EIA_capacity),
                 names_to = "source",
                 values_to = "capacity")
  
  # -------------------------
  # Plot
  # -------------------------
  ggplot(thermal_compare, aes(x = fuel, y = capacity, fill = source)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    theme_minimal() +
    labs(
      title = paste(state_name, "- Thermal Capacity by Fuel (EI (Available) vs EIA)"),
      x = "Fuel Type",
      y = "Capacity (MW)",
      fill = "Source"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 9)
    ) +
    scale_fill_manual(values = c(
      "EI_capacity" = "#FF7F73",
      "EIA_capacity" = "#00C4CC"
    ))
}


```

```{r}
plot_thermal_3bars <- function(ei_df, eia_df, state_name) {
  
  # -------------------------
  # EI Thermal
  # -------------------------
  ei_thermal <- ei_df %>%
    filter(prime_mover_type %in% c("ST", "CT", "CC", "OT")) %>%
    filter(fuel != "NUCLEAR") %>%
    group_by(fuel) %>%
    summarise(
      EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      EI_total_capacity     = sum(rating, na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # EIA Thermal
  # -------------------------
  eia_thermal <- eia_df %>%
    filter(Technology %in% c(
      "Conventional Steam Coal",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Petroleum Liquids",
      "Petroleum Coke",
      "Other Gases"
    )) %>%
    mutate(fuel = case_when(
      Technology == "Conventional Steam Coal" ~ "COAL",
      
      Technology %in% c(
        "Natural Gas Fired Combined Cycle",
        "Natural Gas Steam Turbine",
        "Natural Gas Fired Combustion Turbine",
        "Natural Gas Internal Combustion Engine",
        "Other Gases"
      ) ~ "NATURAL_GAS",
      
      Technology %in% c(
        "Petroleum Liquids",
        "Petroleum Coke"
      ) ~ "DISTILLATE_FUEL_OIL",
      
      TRUE ~ "OTHER"
    )) %>%
    group_by(fuel) %>%
    summarise(
      EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # Combine
  # -------------------------
  thermal_compare <- full_join(ei_thermal, eia_thermal, by = "fuel") %>%
    replace_na(list(
      EI_available_capacity = 0,
      EI_total_capacity = 0,
      EIA_capacity = 0
    )) %>%
    pivot_longer(
      cols = c(EI_available_capacity, EI_total_capacity, EIA_capacity),
      names_to = "source",
      values_to = "capacity"
    )
  
  # -------------------------
  # Plot
  # -------------------------
  ggplot(thermal_compare, aes(x = fuel, y = capacity, fill = source)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    theme_minimal() +
    labs(
      title = paste(state_name, "- Thermal Capacity (EI Available vs Total vs EIA)"),
      x = "Fuel Type",
      y = "Capacity (MW)",
      fill = "Source"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 9)
    ) +
    scale_fill_manual(values = c(
      "EI_available_capacity" = "#FF7F73",   # coral
      "EI_total_capacity"     = "#FFA07A",   # lighter coral
      "EIA_capacity"          = "#00C4CC"    # teal
    ))
}
```



### VA 

```{r run-summary}

va_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "VA Capacity Comparison (Available vs Total EI)")

va_plot_df = va_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )
ggplot(va_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (VA)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )
```

```{r}
plot_thermal_discrepancy(VA_ei_gens, eia_va, "Virginia")
```


```{r}
plot_thermal_3bars(VA_ei_gens, eia_va, "VA")
```

### MD 
```{r }
md_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "MD Capacity Comparison (Available vs Total EI)")
md_plot_df = md_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )

ggplot(md_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (MD)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )

```

```{r}
plot_thermal_discrepancy(MD_ei_gens, eia_md, "Maryland")
```

```{r}
plot_thermal_3bars(MD_ei_gens, eia_md, "MD")
```


### WV 
```{r }
wv_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "WV Capacity Comparison (Available vs Total EI)")
wv_plot_df = wv_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )

ggplot(wv_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (WV)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )

```

```{r}
plot_thermal_discrepancy(WV_ei_gens, eia_wv, "West Virginia")
```

```{r}
plot_thermal_3bars(WV_ei_gens, eia_wv, "WV")
```

---

# EI and EIA comparison

3 sets:

- ‚úÖ Matched (EI ‚ą© EIA) ‚Üí update capacity if needed

- ‚ěē EIA only (EIA ‚ąí EI) ‚Üí ADD to EI

- ‚ěĖ EI only (EI ‚ąí EIA) ‚Üí set unavailable

```{r}
eia_plants = read_xlsx("2___Plant_Y2024.xlsx", skip = 1)
```


```{r}
add_eia_coordinates <- function(eia_df, eia_plants_df, state_name = NULL) {
  
  # ‚ĒÄ‚ĒÄ 0. Standardize join key types ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  eia_df <- eia_df %>% 
    mutate(`Plant Code` = str_trim(as.character(`Plant Code`)))
  
  plants_clean <- eia_plants_df %>%
    mutate(`Plant Code` = str_trim(as.character(`Plant Code`))) %>%
    select(
    `Plant Code`,
    plant_name_ref = `Plant Name`,
    lat = Latitude,
    lon = Longitude
  )
  
  # ‚ĒÄ‚ĒÄ 1. Join ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  merged_df <- eia_df %>% 
    left_join(plants_clean, by = "Plant Code")

      # See if any rows actually joined
      sum(!is.na(merged_df$plant_name_ref))
  
  # ‚ĒÄ‚ĒÄ 2. Validation: Plant Name check ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  name_check <- merged_df %>%
    filter(!is.na(`Plant Name`) & !is.na(plant_name_ref)) %>%
    mutate(name_match = str_trim(tolower(`Plant Name`)) == 
                         str_trim(tolower(plant_name_ref)))
  
  mismatches <- name_check %>% filter(!name_match)
  
  if (nrow(mismatches) > 0) {
    warning(
      paste0(
        "‚ö†ÔłŹ Plant Name mismatches",
        if (!is.null(state_name)) paste0(" (", state_name, ")"),
        ": ", nrow(mismatches), " rows\n",
        paste(head(
          paste0(mismatches$`Plant Name`, " ‚Č† ", mismatches$plant_name_ref),
          5
        ), collapse = "\n")
      )
    )
  }
  
  # ‚ĒÄ‚ĒÄ 3. Missing coordinates check ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  missing_coords <- merged_df %>%
    filter(is.na(lat) | is.na(lon))
  
  if (nrow(missing_coords) > 0) {
    warning(
      paste0(
        "‚ö†ÔłŹ Missing coordinates",
        if (!is.null(state_name)) paste0(" (", state_name, ")"),
        ": ", nrow(missing_coords), " rows"
      )
    )
  }
  
  return(merged_df)
}
```

Add coordinates from the EIA plants file

```{r}
eia_va <- add_eia_coordinates(eia_va, eia_plants, "VA")
eia_md <- add_eia_coordinates(eia_md, eia_plants, "MD")
eia_wv <- add_eia_coordinates(eia_wv, eia_plants, "WV")
```



```{r}



#' Group generators by geographic location and aggregate their capacities
#'
#' @param gens data.frame: Generator dataframe
#' @param lat_col character: Name of latitude column
#' @param lon_col character: Name of longitude column
#' @param capacity_col character: Name of capacity column
#' @param name_col character: Name of generator name column
#'
#' @return data.frame: Grouped generators with aggregated capacity and generator lists
group_generators_by_location = function(gens, lat_col, lon_col, capacity_col, name_col) {
  grouped = gens %>%
    group_by(across(all_of(c(lat_col, lon_col)))) %>%
    summarise(
      gen_names = paste(get(name_col), collapse = "; "),
      n_generators = n(),
      total_capacity = sum(get(capacity_col), na.rm = TRUE),
      first_gen_name = first(get(name_col)),
      .groups = 'drop'
    )
  
  return(grouped)
}

#' Compare generators between EI and EIA datasets using geographic coordinates
#' First groups generators by location, then matches locations
#'
#' @param ei_gens data.frame: EI generators dataframe (must have Lat, Lon, Rating columns)
#' @param eia_gens data.frame: EIA generators dataframe (must have lat, lon, nameplate_capacity columns)
#' @param coord_tolerance numeric: Maximum distance in degrees for matching (default: 0.02 ‚Čą 2.2km)
#'
#' @return list: Contains grouped and detailed views of matches, plus summary statistics
compare_generator_fleets = function(ei_gens, eia_gens, coord_tolerance = 0.2) {
  
  # Step 1: Group generators by location
  cat("\nGrouping generators by location...\n")
  
  # Group EI generators
  
  ei_grouped = ei_gens %>%
    group_by(Lat, Lon) %>%
    summarise(
      GenNames         = paste(GenName, collapse = "; "),
      EI_plant_names = {
  vals <- unique(na.omit(plant_name))
  if (length(vals) == 0) NA_character_ else paste(vals, collapse = "; ")
},  # ūüĎą ADD THIS
      BusNumbers       = paste(BusNumber, collapse = "; "),
      BusNames         = paste(BusName, collapse = "; "),
      n_ei_generators  = n(),
      Rating           = sum(Rating),
      Available        = all(Available),
      Type             = first(Type),
      .groups          = 'drop'
  )
  
  cat("  EI:", nrow(ei_gens), "generators ‚Üí", nrow(ei_grouped), "unique locations\n")
  
  # Group EIA generators
  # Group EIA generators
  eia_grouped = eia_gens %>%
  group_by(lat, lon) %>%
  summarise(
    plant_names = paste(`Plant Name`, collapse = "; "),  # EIA plant names
    plant_codes = paste(`Plant Code`, collapse = "; "),  # EIA plant codes
    generator_ids = paste(`Generator ID`, collapse = "; "),
    n_eia_generators = n(),
    nameplate_capacity = sum(nameplate_capacity, na.rm = TRUE),
    technology = first(Technology),
    .groups = 'drop'
  )
  
  cat("  EIA:", nrow(eia_gens), "generators ‚Üí", nrow(eia_grouped), "unique locations\n")
  
  # Step 2: Match locations
  cat("\nMatching locations (tolerance:", coord_tolerance, "¬į)...\n")
  
  # Function to find closest match
  find_closest_match = function(lat, lon, df, tolerance) {
    distances = sqrt((df$lat - lat)^2 + (df$lon - lon)^2)
    min_dist = min(distances)
    idx = which.min(distances)
    
    if (min_dist <= tolerance) {
      return(idx)
    } else {
      return(NA)
    }
  }
  
  # Match EI locations to EIA locations
  ei_grouped$eia_match_idx = sapply(1:nrow(ei_grouped), function(i) {
    find_closest_match(ei_grouped$Lat[i], ei_grouped$Lon[i], 
                      eia_grouped, coord_tolerance)
  })
  
  # Step 3: Create matched dataset
  in_both_grouped = ei_grouped %>% 
    filter(!is.na(eia_match_idx))
  
  if (nrow(in_both_grouped) > 0) {
    # Add EIA data
    in_both_grouped$eia_lat = eia_grouped$lat[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_lon = eia_grouped$lon[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_plant_names = eia_grouped$plant_names[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_generator_ids = eia_grouped$generator_ids[in_both_grouped$eia_match_idx]
    in_both_grouped$n_eia_generators = eia_grouped$n_eia_generators[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_capacity = eia_grouped$nameplate_capacity[in_both_grouped$eia_match_idx]
    
    # Calculate differences
    in_both_grouped$capacity_diff = in_both_grouped$Rating - in_both_grouped$eia_capacity
    in_both_grouped$capacity_abs_diff = abs(in_both_grouped$capacity_diff)
    
    # Percentage difference with respect to EIA capacity
    in_both_grouped$capacity_pct_diff = (in_both_grouped$Rating - in_both_grouped$eia_capacity) / 
                                         in_both_grouped$eia_capacity * 100
    in_both_grouped$capacity_abs_pct_diff = abs(in_both_grouped$capacity_pct_diff)
    in_both_grouped$capacity_ratio = in_both_grouped$Rating / in_both_grouped$eia_capacity
    
    
    
    # Rename for clarity
    in_both_grouped = in_both_grouped %>%
    rename(
      ei_lat = Lat,
      ei_lon = Lon,
      ei_plant_names = EI_plant_names
    )
}
  
  # Step 4: Create detailed view (expand to individual generators)
  in_both_detailed = data.frame()
  
  if (nrow(in_both_grouped) > 0) {
    for (i in 1:nrow(in_both_grouped)) {
      row = in_both_grouped[i, ]
      
      # Get individual EI generators at this location
      ei_at_location = ei_gens %>%
        filter(Lat == row$ei_lat & Lon == row$ei_lon)
      
      # Add aggregated matching info to each
      ei_at_location$eia_plant_names = row$eia_plant_names
      ei_at_location$eia_generator_ids = row$eia_generator_ids
      ei_at_location$n_eia_generators = row$n_eia_generators
      ei_at_location$n_ei_at_location = row$n_ei_generators
      ei_at_location$eia_capacity = row$eia_capacity
      ei_at_location$total_ei_capacity_at_location = row$Rating
      ei_at_location$eia_lat = row$eia_lat
      ei_at_location$eia_lon = row$eia_lon
      ei_at_location$capacity_diff = row$capacity_diff
      ei_at_location$capacity_abs_diff = row$capacity_abs_diff
      ei_at_location$ei_plant_names = row$ei_plant_names
      
      
      # Rename lat/lon
      ei_at_location = ei_at_location %>%
        rename(ei_lat = Lat, ei_lon = Lon)
      
      in_both_detailed = bind_rows(in_both_detailed, ei_at_location)
    }
  }
  
  # Step 5: Identify unmatched generators
  in_ei_only_grouped = ei_grouped %>% 
    filter(is.na(eia_match_idx))
  
  matched_eia_indices = in_both_grouped$eia_match_idx[!is.na(in_both_grouped$eia_match_idx)]
  in_eia_only_grouped = eia_grouped %>% 
    filter(!(row_number() %in% matched_eia_indices))
  
  # Calculate summary statistics
  total_ei_capacity = sum(ei_gens$Rating, na.rm = TRUE)
  total_eia_capacity = sum(eia_gens$nameplate_capacity, na.rm = TRUE)
  matched_ei_capacity = if(nrow(in_both_grouped) > 0) sum(in_both_grouped$Rating, na.rm = TRUE) else 0
  matched_eia_capacity = if(nrow(in_both_grouped) > 0) 
                        {
                        unique_eia_idx = unique(in_both_grouped$eia_match_idx)
                        matched_eia_capacity = sum(eia_grouped$nameplate_capacity[unique_eia_idx], na.rm = TRUE)
                        } 
                        else {
                        matched_eia_capacity = 0
                        }
  mean_capacity_diff = if(nrow(in_both_grouped) > 0) mean(in_both_grouped$capacity_diff, na.rm = TRUE) else 0
  mean_capacity_abs_diff = if(nrow(in_both_grouped) > 0) mean(in_both_grouped$capacity_abs_diff, na.rm = TRUE) else 0
  
  # Print summary
  cat("\n", paste(rep("=", 60), collapse = ""), "\n")
  cat("Fleet Comparison Results\n")
  cat(paste(rep("=", 60), collapse = ""), "\n")
  cat("Matched locations:", nrow(in_both_grouped), "\n")
  if(nrow(in_both_grouped) > 0) {
    cat("  Total EI generators at matched locations:", sum(in_both_grouped$n_ei_generators), "\n")
    cat("  Total EIA generators at matched locations:", sum(in_both_grouped$n_eia_generators), "\n")
  }
  cat("Unmatched EI locations:", nrow(in_ei_only_grouped), "\n")
  cat("Unmatched EIA locations:", nrow(in_eia_only_grouped), "\n")
  cat("\nCapacity Summary:\n")
  cat("  Total EI capacity:", round(total_ei_capacity, 2), "MW\n")
  cat("  Total EIA capacity:", round(total_eia_capacity, 2), "MW\n")
  cat("  Matched EI capacity:", round(matched_ei_capacity, 2), "MW\n")
  cat("  Matched EIA capacity:", round(matched_eia_capacity, 2), "MW\n")
  cat("  Mean capacity difference:", round(mean_capacity_diff, 2), "MW\n")
  cat("  Mean absolute capacity difference:", round(mean_capacity_abs_diff, 2), "MW\n")
  cat(paste(rep("=", 60), collapse = ""), "\n\n")
  
  # Return results as a list
  return(list(
    in_both_grouped = in_both_grouped,
    in_both_detailed = in_both_detailed,
    in_ei_only_grouped = in_ei_only_grouped,
    in_eia_only_grouped = in_eia_only_grouped,
    ei_grouped = ei_grouped,
    eia_grouped = eia_grouped,
    summary = list(
      total_ei_capacity = total_ei_capacity,
      total_eia_capacity = total_eia_capacity,
      matched_ei_capacity = matched_ei_capacity,
      matched_eia_capacity = matched_eia_capacity,
      mean_capacity_diff = mean_capacity_diff,
      mean_capacity_abs_diff = mean_capacity_abs_diff,
      n_matched_locations = nrow(in_both_grouped),
      n_ei_only_locations = nrow(in_ei_only_grouped),
      n_eia_only_locations = nrow(in_eia_only_grouped)
    )
  ))
}


```

```{r}
#' Compare generator fleets for a specific technology type
#'
#' @param ei_gens data.frame: Full EI generators dataframe (virginia_resources.csv)
#' @param eia_gens data.frame: Full EIA generators dataframe
#' @param type character: "Solar", "Wind", "Hydro", "Thermal", "Storage"
#' @param coord_tolerance numeric: tolerance in degrees (default 0.2)
compare_by_type = function(ei_gens, eia_gens, type = "Solar", coord_tolerance = 0.2,
                           eia_type_col = "Technology") {
  
  # ‚ĒÄ‚ĒÄ 1. Type ‚Üí filter mappings ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  ei_type_map = list(
  "Solar"   = c("Solar"),
  "Wind"    = c("Wind"),
  "Hydro"   = c("Hydro"),
  "Pumped Storage" = c("Pumped Storage"),   # ūüĎą ADD THIS
  "Thermal" = c("ThermalStandard", "ThermalMultiStart"),
  "Nuclear" = c("ThermalStandard"),
  "Storage" = c("Storage")
)

eia_type_map = list(
  "Solar"   = c("Solar Photovoltaic"),
  "Wind"    = c("Onshore Wind Turbine", "Offshore Wind Turbine"),
  "Hydro"   = c("Conventional Hydroelectric", "Run-of-River Hydroelectric"),
  "Pumped Storage" = c("Hydroelectric Pumped Storage"),  # ūüĎą ADD THIS
  "Thermal" = c("Natural Gas Steam Turbine", "Natural Gas Fired Combined Cycle",
                "Natural Gas Fired Combustion Turbine", "Coal Steam Turbine"),
  "Nuclear" = c("Nuclear"),
  "Storage" = c("Batteries")
)
  
  if (!type %in% names(ei_type_map)) {
    stop("Invalid type. Choose from: ", paste(names(ei_type_map), collapse = ", "))
  }
  
  # ‚ĒÄ‚ĒÄ 2. Filter by type ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  ei_filtered = ei_gens %>%
  filter(
    generator_type %in% ei_type_map[[type]] &
    !(type == "Thermal" & fuel == "NUCLEAR")
  )
  
  # Use column override if needed
  if (!eia_type_col %in% colnames(eia_gens)) {
    stop(paste("EIA type column", eia_type_col, "not found in EIA dataset"))
  }
  
  # eia_filtered = eia_gens %>% filter(.data[[eia_type_col]] %in% eia_type_map[[type]])
  eia_filtered = eia_gens %>% filter(tech_group == type)
  
  cat("=== Type:", type, "===\n")
  cat("EI generators:  ", nrow(ei_filtered), "\n")
  cat("EIA generators: ", nrow(eia_filtered), "\n")
  
  if (nrow(ei_filtered) == 0 || nrow(eia_filtered) == 0) {
    warning("No generators found for type: ", type)
    return(NULL)
  }
  
  # ‚ĒÄ‚ĒÄ 3. Rename EI columns to match compare_generator_fleets expectations ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  ei_renamed = ei_filtered %>%
    rename(
      GenName   = name,
      BusNumber = bus_number,
      BusName   = bus,
      Lat       = lat,
      Lon       = lon,
      Rating    = rating,
      Available = available,
      Type      = generator_type
    )
  
  # ‚ĒÄ‚ĒÄ 4. Run comparison ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ‚ĒÄ
  result = compare_generator_fleets(
    ei_gens        = ei_renamed,
    eia_gens       = eia_filtered,
    coord_tolerance = coord_tolerance
  )
  
  return(result)
}
```

```{r}
clean_name <- function(x) {
  x %>%
    tolower() %>%
    stringr::str_replace_all("[^a-z0-9 ]", " ") %>%  # remove symbols like _
    stringr::str_squish()
}



compute_name_similarity <- function(eia_name, bus_name) {
  eia_clean <- clean_name(eia_name)
  bus_clean <- clean_name(bus_name)
  
  # Jaro-Winkler similarity (0 to 1, higher = better)
  jw <- stringsim(eia_clean, bus_clean, method = "jw")
  
  # Check if any word overlaps
  eia_words <- unlist(strsplit(eia_clean, " "))
  bus_words <- unlist(strsplit(bus_clean, " "))
  
  word_overlap <- length(intersect(eia_words, bus_words)) / length(unique(eia_words))
  
  return(list(
    jw_similarity = jw,
    word_overlap = word_overlap
  ))
}



get_best_similarity <- function(eia_names, bus_names) {
  
  eia_list <- strsplit(eia_names, ";\\s*")[[1]]
  bus_list <- strsplit(bus_names, ";\\s*")[[1]]
  
  scores <- expand.grid(eia = eia_list, bus = bus_list, stringsAsFactors = FALSE) %>%
    rowwise() %>%
    mutate(
      sim = list(compute_name_similarity(eia, bus)),
      jw = sim$jw_similarity,
      overlap = sim$word_overlap
    ) %>%
    ungroup()
  
  # Take best match
  best <- scores %>%
    arrange(desc(jw), desc(overlap)) %>%
    slice(1)
  
  return(best[, c("jw", "overlap")])
}
```


## VA

```{r}
compare_nonthermal <- function(ei_df, eia_df, type_name, available_only = TRUE, prime_mover_filter = NULL) {
  # -------------------------
  # Filter EI
  # -------------------------
  ei_filtered <- ei_df
  if (!is.null(prime_mover_filter)) {
    ei_filtered <- ei_filtered %>% filter(prime_mover_type %in% prime_mover_filter)
  }
  if (available_only) {
    ei_filtered <- ei_filtered %>% filter(available == TRUE)
  }

  # Filter EIA
  eia_filtered <- eia_df %>% filter(tech_group == type_name)
  if (!is.null(prime_mover_filter) && type_name == "Hydro") {
    # special mapping for Pumped Storage
    eia_filtered <- eia_filtered %>% mutate(tech_group = "Hydro")
  }

  # -------------------------
  # Compare
  # -------------------------
  results <- compare_by_type(ei_filtered, eia_filtered, type = type_name, coord_tolerance = 0.5)

  # -------------------------
  # Handle empty results
  # -------------------------
  if (is.null(results$in_both_grouped) || nrow(results$in_both_grouped) == 0) {
    warning(paste("No matched generators for type:", type_name))
    summary_results <- tibble()
  } else {
    summary_results <- results$in_both_grouped %>%
      rowwise() %>%
      mutate(
        sim = list(get_best_similarity(eia_plant_names, BusNames)),
        name_similarity = sim$jw,
        word_overlap = sim$overlap
      ) %>%
      ungroup() %>%
      mutate(
        eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
        bus_names_short = stringr::str_trunc(BusNames, width = 40)
      ) %>%
      dplyr::select(
        ei_plant_names,
        eia_plant_names_short,
        bus_names_short,
        capacity_abs_pct_diff,
        name_similarity,
        capacity_diff,
        Rating,
        eia_capacity,
        n_ei_generators,
        n_eia_generators,
        GenNames,
        capacity_abs_diff,
        capacity_pct_diff,
        ei_lat, eia_lat, ei_lon, eia_lon
      ) %>%
      dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity))
  }

  return(list(results = results, summary = summary_results))
}
```



### Solar

#### All EI


```{r}
solar_va_all <- compare_nonthermal(VA_ei_gens, eia_va, "Solar", available_only = FALSE)
solar_va_all$summary
```

#### EI available only
```{r}
solar_va_avail <- compare_nonthermal(VA_ei_gens, eia_va, "Solar", available_only = TRUE)
solar_va_avail$summary
```

### Wind

#### AL EI

#### All EI


```{r}
wind_va_all <- compare_nonthermal(VA_ei_gens, eia_va, "Wind", available_only = FALSE)
wind_va_all$summary
```

#### EI available only
```{r}
wind_va_avail <- compare_nonthermal(VA_ei_gens, eia_va, "Wind", available_only = TRUE)
wind_va_avail$summary
```


### Hydro

#### EI all

```{r}
hydro_va_all <- compare_nonthermal(VA_ei_gens, eia_va, "Hydro", available_only = FALSE, prime_mover_filter = "HY")
hydro_va_all$summary
# eia_va_hydro = eia_va %>% filter(tech_group == "Hydro")
# #eia_va_hydro_tech = eia_va_hydro %>% mutate(technology = "Hydro")
# hydro_results = compare_by_type(VA_ei_gens[VA_ei_gens$prime_mover_type == "HY",], eia_va_hydro, type = "Hydro", coord_tolerance = 0.5)
# 
# hydro_results$in_both_grouped <- hydro_results$in_both_grouped %>%
#   rowwise() %>%
#   mutate(
#     sim = list(get_best_similarity(eia_plant_names, BusNames)),
#     name_similarity = sim$jw,
#     word_overlap = sim$overlap
#   ) %>%
#   ungroup()
# 
# # Check for matches
# if (!is.null(hydro_results) && !is.null(hydro_results$in_both_grouped)) {
#   hydro_results$in_both_grouped %>%
#     mutate(
#     eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
#     bus_names_short = stringr::str_trunc(BusNames, width = 40),
#   ) %>%
#   dplyr::select(
#     ei_plant_names,
#     eia_plant_names_short,
#     bus_names_short,
#     capacity_abs_pct_diff,
#     name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
#     capacity_diff,
#     Rating,
#     eia_capacity,
#     n_ei_generators,
#     n_eia_generators,
#     GenNames,                # ūüĎą EI generator names
#     capacity_abs_diff,
#     capacity_pct_diff,
#     ei_lat, eia_lat, ei_lon, eia_lon
#   ) %>%
#   dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
# } else {
#   cat("No matched Hydro generators found.\n")
# }
```

#### EI available only


```{r}
hydro_va_available <- compare_nonthermal(VA_ei_gens, eia_va, "Hydro", available_only = TRUE, prime_mover_filter = "HY")
hydro_va_available$summary

# eia_va_hydro = eia_va %>% filter(tech_group == "Hydro")
# #eia_va_hydro_tech = eia_va_hydro %>% mutate(technology = "Hydro")
# hydro_results = compare_by_type(VA_ei_gens[(VA_ei_gens$prime_mover_type == "HY" & VA_ei_gens$available==TRUE),], eia_va_hydro, type = "Hydro", coord_tolerance = 0.5)
# 
# hydro_results$in_both_grouped <- hydro_results$in_both_grouped %>%
#   rowwise() %>%
#   mutate(
#     sim = list(get_best_similarity(eia_plant_names, BusNames)),
#     name_similarity = sim$jw,
#     word_overlap = sim$overlap
#   ) %>%
#   ungroup()
# 
# # Check for matches
# if (!is.null(hydro_results) && !is.null(hydro_results$in_both_grouped)) {
#   hydro_results$in_both_grouped %>%
#     mutate(
#     eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
#     bus_names_short = stringr::str_trunc(BusNames, width = 40),
#   ) %>%
#   dplyr::select(
#     ei_plant_names,
#     eia_plant_names_short,
#     bus_names_short,
#     capacity_abs_pct_diff,
#     name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
#     capacity_diff,
#     Rating,
#     eia_capacity,
#     n_ei_generators,
#     n_eia_generators,
#     GenNames,                # ūüĎą EI generator names
#     capacity_abs_diff,
#     capacity_pct_diff,
#     ei_lat, eia_lat, ei_lon, eia_lon
#   ) %>%
#   dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
# } else {
#   cat("No matched Hydro generators found.\n")
# }
```

### Pumped Storage

#### EI all

```{r}

# Pumped Storage comparison
ei_va_ps <- VA_ei_gens %>% filter(prime_mover_type == "PS") %>%
  mutate(generator_type = "Hydro")  # temporarily trick type mapping

eia_va_ps <- eia_va %>% filter(tech_group == "Pumped Storage") %>%
  mutate(tech_group = "Hydro")      # temporarily trick type mapping

ps_results <- compare_by_type(
  ei_va_ps,
  eia_va_ps,
  type = "Hydro",         # keep Hydro to match your function
  coord_tolerance = 0.5
)


ps_results$in_both_grouped <- ps_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

# Check for matches
if (!is.null(ps_results) && !is.null(ps_results$in_both_grouped)) {
  ps_results$in_both_grouped %>%
    mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
} else {
  cat("No matched Pumped Storage generators found.\n")
}
```

#### EI available only

```{r}

# Pumped Storage comparison
ei_va_ps <- VA_ei_gens %>% filter(prime_mover_type == "PS", available==TRUE) %>%
  mutate(generator_type = "Hydro")  # temporarily trick type mapping

eia_va_ps <- eia_va %>% filter(tech_group == "Pumped Storage") %>%
  mutate(tech_group = "Hydro")      # temporarily trick type mapping

ps_results <- compare_by_type(
  ei_va_ps,
  eia_va_ps,
  type = "Hydro",         # keep Hydro to match your function
  coord_tolerance = 0.5
)


ps_results$in_both_grouped <- ps_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

# Check for matches
if (!is.null(ps_results) && !is.null(ps_results$in_both_grouped)) {
  ps_results$in_both_grouped %>%
    mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
} else {
  cat("No matched Pumped Storage generators found.\n")
}
```



### Thermal

#### EI all

```{r}
# Map technology for EI and EIA
ei_va_mapped  <- map_technology(VA_ei_gens, "EI")
eia_va_mapped <- map_technology(eia_va, "EIA")

# Filter EI available if needed
ei_va_thermal <- ei_va_mapped %>% filter(tech_group == "Thermal")
ei_va_thermal_available <- ei_va_mapped %>% filter(tech_group == "Thermal", available == TRUE)

# Filter EIA thermal
eia_va_thermal <- eia_va_mapped %>% filter(tech_group == "Thermal")



# Compare all thermal
thermal_results <- compare_by_type(
  ei_va_thermal,
  eia_va_thermal,
  type = "Thermal",
  coord_tolerance = 0.5
)

thermal_results$in_both_grouped <- thermal_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

thermal_results$in_both_grouped %>%
  mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )


```

#### EI available only

```{r}

# Compare available only
thermal_results <- compare_by_type(
  ei_va_thermal_available,
  eia_va_thermal,
  type = "Thermal",
  coord_tolerance = 0.5
)

thermal_results$in_both_grouped <- thermal_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

thermal_results$in_both_grouped %>%
  mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )


```

### Thermal by fuel

```{r}
compare_thermal_by_fuel <- function(ei_df, eia_df, fuel_type, available_only = TRUE) {
  # -------------------------
  # Filter EI
  # -------------------------
  ei_filtered <- ei_df %>%
    filter(fuel == fuel_type) %>%
    { if (available_only) filter(., available == TRUE) else . }

  # -------------------------
  # Filter EIA
  # -------------------------
  eia_filtered <- switch(fuel_type,
    "NATURAL_GAS" = eia_df %>% filter(Technology %in% c(
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Other Gases"
    )),
    "COAL" = eia_df %>% filter(Technology == "Conventional Steam Coal"),
    "DISTILLATE_FUEL_OIL" = eia_df %>% filter(Technology %in% c("Petroleum Liquids", "Petroleum Coke")),
    # default: all others
    eia_df %>% filter(!Technology %in% c(
      "Conventional Steam Coal",
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Petroleum Liquids",
      "Petroleum Coke"
    ))
  )

  # -------------------------
  # Compare
  # -------------------------
  results <- compare_by_type(ei_filtered, eia_filtered, type = "Thermal", coord_tolerance = 0.5)

  # -------------------------
  # Check for empty results
  # -------------------------
  if (is.null(results$in_both_grouped) || nrow(results$in_both_grouped) == 0) {
    warning(paste("No matched generators for fuel:", fuel_type))
    summary_results <- tibble()  # return empty tibble
  } else {
    summary_results <- results$in_both_grouped %>%
      rowwise() %>%
      mutate(
        sim = list(get_best_similarity(eia_plant_names, BusNames)),
        name_similarity = sim$jw,
        word_overlap = sim$overlap
      ) %>%
      ungroup() %>%
      mutate(
        eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
        bus_names_short = stringr::str_trunc(BusNames, width = 40)
      ) %>%
      dplyr::select(
        ei_plant_names,
        eia_plant_names_short,
        bus_names_short,
        capacity_abs_pct_diff,
        name_similarity,
        capacity_diff,
        Rating,
        eia_capacity,
        n_ei_generators,
        n_eia_generators,
        GenNames,
        capacity_abs_diff,
        capacity_pct_diff,
        ei_lat, eia_lat, ei_lon, eia_lon
      ) %>%
      dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity))
  }

  return(list(results = results, summary = summary_results))
}
```


#### Natural Gas

##### EI all

```{r}
# Natural Gas, all EI
ng_all <- compare_thermal_by_fuel(ei_va_thermal, eia_va_thermal, "NATURAL_GAS", available_only = FALSE)
ng_all$summary

```

##### EI available only

```{r}
# Natural Gas, available only
ng_avail <- compare_thermal_by_fuel(ei_va_thermal, eia_va_thermal, "NATURAL_GAS", available_only = TRUE)
ng_avail$summary
```

#### Coal

##### EI all

```{r}
# Coal, available only
coal_all = compare_thermal_by_fuel(ei_va_thermal, eia_va_thermal, "COAL", available_only = FALSE)
coal_all$summary 
```

##### EI available

```{r}
# Coal, available only
coal_all = compare_thermal_by_fuel(ei_va_thermal, eia_va_thermal, "COAL", available_only = TRUE)
#coal_all$summary 
```

#### Distillate Fuel Oil

##### All EI

```{r}
dist_all <- compare_thermal_by_fuel(ei_va_thermal, eia_va_thermal, "DISTILLATE_FUEL_OIL", available_only = FALSE)
dist_all$summary 
```

##### EI available only

```{r}
dist_all <- compare_thermal_by_fuel(ei_va_thermal, eia_va_thermal, "DISTILLATE_FUEL_OIL", available_only = TRUE)
dist_all$summary 
```

## MD

### Solar

#### All EI


```{r}
solar_md_all <- compare_nonthermal(MD_ei_gens, eia_md, "Solar", available_only = FALSE)
solar_md_all$summary
```

#### EI available only
```{r}
solar_md_avail <- compare_nonthermal(MD_ei_gens, eia_md, "Solar", available_only = TRUE)
solar_md_avail$summary
```

### Wind

#### AL EI

#### All EI


```{r}
wind_md_all <- compare_nonthermal(MD_ei_gens, eia_md, "Wind", available_only = FALSE)
wind_md_all$summary
```

#### EI available only
```{r}
wind_md_avail <- compare_nonthermal(MD_ei_gens, eia_md, "Wind", available_only = TRUE)
wind_md_avail$summary
```


### Hydro

#### EI all

```{r}
hydro_md_all <- compare_nonthermal(MD_ei_gens, eia_md, "Hydro", available_only = FALSE, prime_mover_filter = "HY")
hydro_md_all$summary

```

#### EI available only


```{r}
hydro_md_available <- compare_nonthermal(MD_ei_gens, eia_md, "Hydro", available_only = TRUE, prime_mover_filter = "HY")
hydro_md_available$summary
```

### Pumped Storage

#### EI all

```{r}

# Pumped Storage comparison
ei_md_ps <- MD_ei_gens %>% filter(prime_mover_type == "PS") %>%
  mutate(generator_type = "Hydro")  # temporarily trick type mapping

eia_md_ps <- eia_md %>% filter(tech_group == "Pumped Storage") %>%
  mutate(tech_group = "Hydro")      # temporarily trick type mapping

ps_results <- compare_by_type(
  ei_md_ps,
  eia_md_ps,
  type = "Hydro",         # keep Hydro to match your function
  coord_tolerance = 0.5
)


# ps_results$in_both_grouped <- ps_results$in_both_grouped %>%
#   rowwise() %>%
#   mutate(
#     sim = list(get_best_similarity(eia_plant_names, BusNames)),
#     name_similarity = sim$jw,
#     word_overlap = sim$overlap
#   ) %>%
#   ungroup()

# Check for matches
if (!is.null(ps_results) && !is.null(ps_results$in_both_grouped)) {
  ps_results$in_both_grouped %>%
    mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
} else {
  cat("No matched Pumped Storage generators found.\n")
}
```

#### EI available only

```{r}

# Pumped Storage comparison
ei_md_ps <- MD_ei_gens %>% filter(prime_mover_type == "PS", available==TRUE) %>%
  mutate(generator_type = "Hydro")  # temporarily trick type mapping

eia_md_ps <- eia_md %>% filter(tech_group == "Pumped Storage") %>%
  mutate(tech_group = "Hydro")      # temporarily trick type mapping

ps_results <- compare_by_type(
  ei_md_ps,
  eia_md_ps,
  type = "Hydro",         # keep Hydro to match your function
  coord_tolerance = 0.5
)


# ps_results$in_both_grouped <- ps_results$in_both_grouped %>%
#   rowwise() %>%
#   mutate(
#     sim = list(get_best_similarity(eia_plant_names, BusNames)),
#     name_similarity = sim$jw,
#     word_overlap = sim$overlap
#   ) %>%
#   ungroup()

# Check for matches
if (!is.null(ps_results) && !is.null(ps_results$in_both_grouped)) {
  ps_results$in_both_grouped %>%
    mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
} else {
  cat("No matched Pumped Storage generators found.\n")
}
```



### Thermal

#### EI all

```{r}
# Map technology for EI and EIA
ei_md_mapped  <- map_technology(MD_ei_gens, "EI")
eia_md_mapped <- map_technology(eia_md, "EIA")

# Filter EI available if needed
ei_md_thermal <- ei_md_mapped %>% filter(tech_group == "Thermal")
ei_md_thermal_available <- ei_md_mapped %>% filter(tech_group == "Thermal", available == TRUE)

# Filter EIA thermal
eia_md_thermal <- eia_md_mapped %>% filter(tech_group == "Thermal")



# Compare all thermal
thermal_results <- compare_by_type(
  ei_md_thermal,
  eia_md_thermal,
  type = "Thermal",
  coord_tolerance = 0.5
)

thermal_results$in_both_grouped <- thermal_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

thermal_results$in_both_grouped %>%
  mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )


```

#### EI available only

```{r}

# Compare available only
thermal_results <- compare_by_type(
  ei_md_thermal_available,
  eia_md_thermal,
  type = "Thermal",
  coord_tolerance = 0.5
)

thermal_results$in_both_grouped <- thermal_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

thermal_results$in_both_grouped %>%
  mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )


```

### Thermal by fuel


#### Natural Gas

##### EI all

```{r}
# Natural Gas, all EI
ng_all <- compare_thermal_by_fuel(ei_md_thermal, eia_md_thermal, "NATURAL_GAS", available_only = FALSE)
ng_all$summary

```

##### EI available only

```{r}
# Natural Gas, available only
ng_avail <- compare_thermal_by_fuel(ei_md_thermal, eia_md_thermal, "NATURAL_GAS", available_only = TRUE)
ng_avail$summary
```

#### Coal

##### EI all

```{r}
# Coal, available only
coal_all = compare_thermal_by_fuel(ei_md_thermal, eia_md_thermal, "COAL", available_only = FALSE)
coal_all$summary 
```

##### EI available

```{r}
# Coal, available only
coal_all = compare_thermal_by_fuel(ei_md_thermal, eia_md_thermal, "COAL", available_only = TRUE)
#coal_all$summary 
```

#### Distillate Fuel Oil

##### All EI

```{r}
dist_all <- compare_thermal_by_fuel(ei_md_thermal, eia_md_thermal, "DISTILLATE_FUEL_OIL", available_only = FALSE)
dist_all$summary 
```

##### EI available only

```{r}
dist_all <- compare_thermal_by_fuel(ei_md_thermal, eia_md_thermal, "DISTILLATE_FUEL_OIL", available_only = TRUE)
dist_all$summary 
```


## WV

### Solar

#### All EI


```{r}
solar_wv_all <- compare_nonthermal(WV_ei_gens, eia_wv, "Solar", available_only = FALSE)
solar_wv_all$summary
```

#### EI available only
```{r}
solar_wv_avail <- compare_nonthermal(WV_ei_gens, eia_wv, "Solar", available_only = TRUE)
solar_wv_avail$summary
```

### Wind

#### AL EI

#### All EI


```{r}
wind_wv_all <- compare_nonthermal(WV_ei_gens, eia_wv, "Wind", available_only = FALSE)
wind_wv_all$summary
```

#### EI available only
```{r}
wind_wv_avail <- compare_nonthermal(WV_ei_gens, eia_wv, "Wind", available_only = TRUE)
wind_wv_avail$summary
```


### Hydro

#### EI all

```{r}
hydro_wv_all <- compare_nonthermal(WV_ei_gens, eia_wv, "Hydro", available_only = FALSE, prime_mover_filter = "HY")
hydro_wv_all$summary

```

#### EI available only


```{r}
hydro_wv_available <- compare_nonthermal(WV_ei_gens, eia_wv, "Hydro", available_only = TRUE, prime_mover_filter = "HY")
hydro_wv_available$summary
```

### Pumped Storage

#### EI all

```{r}

# Pumped Storage comparison
ei_wv_ps <- WV_ei_gens %>% filter(prime_mover_type == "PS") %>%
  mutate(generator_type = "Hydro")  # temporarily trick type mapping

eia_wv_ps <- eia_wv %>% filter(tech_group == "Pumped Storage") %>%
  mutate(tech_group = "Hydro")      # temporarily trick type mapping

ps_results <- compare_by_type(
  ei_wv_ps,
  eia_wv_ps,
  type = "Hydro",         # keep Hydro to match your function
  coord_tolerance = 0.5
)


# ps_results$in_both_grouped <- ps_results$in_both_grouped %>%
#   rowwise() %>%
#   mutate(
#     sim = list(get_best_similarity(eia_plant_names, BusNames)),
#     name_similarity = sim$jw,
#     word_overlap = sim$overlap
#   ) %>%
#   ungroup()

# Check for matches
if (!is.null(ps_results) && !is.null(ps_results$in_both_grouped)) {
  ps_results$in_both_grouped %>%
    mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
} else {
  cat("No matched Pumped Storage generators found.\n")
}
```

#### EI available only

```{r}

# Pumped Storage comparison
ei_wv_ps <- WV_ei_gens %>% filter(prime_mover_type == "PS", available==TRUE) %>%
  mutate(generator_type = "Hydro")  # temporarily trick type mapping

eia_wv_ps <- eia_wv %>% filter(tech_group == "Pumped Storage") %>%
  mutate(tech_group = "Hydro")      # temporarily trick type mapping

ps_results <- compare_by_type(
  ei_wv_ps,
  eia_wv_ps,
  type = "Hydro",         # keep Hydro to match your function
  coord_tolerance = 0.5
)


# ps_results$in_both_grouped <- ps_results$in_both_grouped %>%
#   rowwise() %>%
#   mutate(
#     sim = list(get_best_similarity(eia_plant_names, BusNames)),
#     name_similarity = sim$jw,
#     word_overlap = sim$overlap
#   ) %>%
#   ungroup()

# Check for matches
if (!is.null(ps_results) && !is.null(ps_results$in_both_grouped)) {
  ps_results$in_both_grouped %>%
    mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )
} else {
  cat("No matched Pumped Storage generators found.\n")
}
```



### Thermal

#### EI all

```{r}
# Map technology for EI and EIA
ei_wv_mapped  <- map_technology(WV_ei_gens, "EI")
eia_wv_mapped <- map_technology(eia_wv, "EIA")

# Filter EI available if needed
ei_wv_thermal <- ei_wv_mapped %>% filter(tech_group == "Thermal")
ei_wv_thermal_available <- ei_wv_mapped %>% filter(tech_group == "Thermal", available == TRUE)

# Filter EIA thermal
eia_wv_thermal <- eia_wv_mapped %>% filter(tech_group == "Thermal")



# Compare all thermal
thermal_results <- compare_by_type(
  ei_wv_thermal,
  eia_wv_thermal,
  type = "Thermal",
  coord_tolerance = 0.5
)

thermal_results$in_both_grouped <- thermal_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

thermal_results$in_both_grouped %>%
  mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )


```

#### EI available only

```{r}

# Compare available only
thermal_results <- compare_by_type(
  ei_wv_thermal_available,
  eia_wv_thermal,
  type = "Thermal",
  coord_tolerance = 0.5
)

thermal_results$in_both_grouped <- thermal_results$in_both_grouped %>%
  rowwise() %>%
  mutate(
    sim = list(get_best_similarity(eia_plant_names, BusNames)),
    name_similarity = sim$jw,
    word_overlap = sim$overlap
  ) %>%
  ungroup()

thermal_results$in_both_grouped %>%
  mutate(
    eia_plant_names_short = stringr::str_trunc(eia_plant_names, width = 30),
    bus_names_short = stringr::str_trunc(BusNames, width = 40),
  ) %>%
  dplyr::select(
    ei_plant_names,
    eia_plant_names_short,
    bus_names_short,
    capacity_abs_pct_diff,
    name_similarity,         # ūüĎą similarity score (Jaro-Winkler)
    capacity_diff,
    Rating,
    eia_capacity,
    n_ei_generators,
    n_eia_generators,
    GenNames,                # ūüĎą EI generator names
    capacity_abs_diff,
    capacity_pct_diff,
    ei_lat, eia_lat, ei_lon, eia_lon
  ) %>%
  dplyr::arrange(capacity_abs_pct_diff, desc(name_similarity) )


```

### Thermal by fuel


#### Natural Gas

##### EI all

```{r}
# Natural Gas, all EI
ng_all <- compare_thermal_by_fuel(ei_wv_thermal, eia_wv_thermal, "NATURAL_GAS", available_only = FALSE)
ng_all$summary

```

##### EI available only

```{r}
# Natural Gas, available only
ng_avail <- compare_thermal_by_fuel(ei_wv_thermal, eia_wv_thermal, "NATURAL_GAS", available_only = TRUE)
ng_avail$summary
```

#### Coal

##### EI all

```{r}
# Coal, available only
coal_all = compare_thermal_by_fuel(ei_wv_thermal, eia_wv_thermal, "COAL", available_only = FALSE)
coal_all$summary 
```

##### EI available

```{r}
# Coal, available only
coal_all = compare_thermal_by_fuel(ei_wv_thermal, eia_wv_thermal, "COAL", available_only = TRUE)
#coal_all$summary 
```

#### Distillate Fuel Oil

##### All EI

```{r}
dist_all <- compare_thermal_by_fuel(ei_wv_thermal, eia_wv_thermal, "DISTILLATE_FUEL_OIL", available_only = FALSE)
dist_all$summary 
```

##### EI available only

```{r}
dist_all <- compare_thermal_by_fuel(ei_wv_thermal, eia_wv_thermal, "DISTILLATE_FUEL_OIL", available_only = TRUE)
dist_all$summary 
```




