---
title: "Data Cleaning & Alignment (EI/Sienna ↔ EIA)"
author: "Sara Abril-Guevara"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: true
    toc_float: true
    df_print: paged
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) 
library(readr) 
library(ggplot2) 
library(tidyr) 
library(knitr)
library(readxl)
```

# Overview

This document summarizes the data cleaning and validation process used to align generator fleets from EI/Sienna and EIA datasets for selected states (MN, IL, MD, VA, WV).

The goal is to ensure consistency in installed capacity before running simulation models.


## Methodology

- DONE: Extract EI/Sienna generators and bus attributes using Kestrel

- DONE: Aggregate generators by geographic coordinates

- DONE: Standardize technology categories across datasets

- WIP: Compare installed capacity by state and technology

- WIP: Flag differences for further spatial matching and manual review

```{r cars, message = FALSE}
# EI datasets 
VA_ei_gens = read_csv("~/Downloads/virginia_resources.csv") 
MD_ei_gens = read_csv("~/Downloads/maryland_resources.csv") 

#WV_ei_gens = read_csv("~/Downloads/westvirginia_resources.csv") 
# EIA datasets 
# Renewable 
eia_va_renew = read_csv("~/Documents/GitHub/va_updates/renewable_VA_EIA.csv") 
# Hydro 
eia_va_hydro = read_csv("~/Documents/GitHub/va_updates/va_hydro_EIA.csv") 
# Thermal 
eia_va_thermal = read_excel("~/Documents/GitHub/va_updates/Thermal_EIA.xlsx")

# Combine all VA EIA datasets 
eia_va <- bind_rows( eia_va_renew, eia_va_hydro, eia_va_thermal )

# Renewable 
eia_md_renew = read_csv("~/Documents/GitHub/md_update/Renewable_EIA.csv") 
# Hydro 
eia_md_hydro = read_csv("~/Documents/GitHub/md_update/Hydro_EIA.csv")
# Thermal 
eia_md_thermal = read_csv("~/Documents/GitHub/md_update/Thermal_EIA.csv")
eia_md_renew <- eia_md_renew %>%
  mutate(`generator ID` = as.character(`generator ID`))

eia_md_hydro <- eia_md_hydro %>%
  mutate(`generator ID` = as.character(`generator ID`))

eia_md_thermal <- eia_md_thermal %>%
  mutate(`generator ID` = as.character(`generator ID`))

eia_md <- bind_rows( eia_md_renew, eia_md_hydro, eia_md_thermal )
#eia_wv = read_csv("renewable_WV_EIA.csv")
```

```{r}
clean_eia <- function(df, state_name) {
  df %>%
    mutate(
      # Fill technology using mode when missing
      technology = case_when(
        !is.na(technology) ~ technology,
        mode == "Storage" ~ "Pumped Storage Hydroelectric",
        mode == "RoR" ~ "Run-of-River Hydroelectric",
        TRUE ~ "Unknown"
      ),
      
      # Add state
      state = state_name
    )
}
eia_va <- clean_eia(eia_va, "VA")
eia_md <- clean_eia(eia_md, "MD")

clean_eia_final <- function(df) {
  df %>%
    select(
      plant,
      `generator ID`,
      technology,
      lat,
      lon,
      nameplate_capacity,
      ba,
      state
    ) %>%
    filter(!is.na(lat) & !is.na(lon))  # remove bad rows
}

eia_va <- clean_eia_final(eia_va)
eia_md <- clean_eia_final(eia_md)
```



## Technology Mapping

```{r tech-mapping}
map_technology = function(df, source = "EI") {
  
  if (source == "EI") {
    df %>%
      mutate(tech_group = case_when(
        fuel == "NUCLEAR" ~ "Nuclear",
        generator_type %in% c("Solar") ~ "Solar",
        generator_type %in% c("Wind") ~ "Wind",
        generator_type %in% c("Hydro") ~ "Hydro",
        generator_type %in% c("ThermalStandard", "ThermalMultiStart") ~ "Thermal",
        generator_type %in% c("Storage") ~ "Storage",
        TRUE ~ "Other"
      ))
    
  } else {
    df %>%
      mutate(tech_group = case_when(
        grepl("Solar", technology) ~ "Solar",
        grepl("Wind", technology) ~ "Wind",
        grepl("Hydro", technology) ~ "Hydro",
        grepl("Pumped Storage", technology) ~ "Storage",
        grepl("Battery|Storage", technology) ~ "Storage",
        grepl("Nuclear", technology) ~ "Nuclear",
        grepl("Gas|Coal|Petroleum|Steam|Combustion", technology) ~ "Thermal",
        grepl("Biomass|Waste|Landfill", technology) ~ "Other",
        TRUE ~ "Other"
      ))
  }
}

# Summarize function including units for both EI and EIA
summarize_capacity_units <- function(ei_df, eia_df, state_name) {
  
  ei_clean  <- map_technology(ei_df, "EI")
  eia_clean <- map_technology(eia_df, "EIA")
  
  # EI summary
  ei_summary <- ei_clean %>%
    group_by(tech_group) %>%
    summarise(
      EI_total_units = n(),
      EI_available_units = sum(available == TRUE, na.rm = TRUE),
      EI_total_capacity = sum(rating, na.rm = TRUE),
      EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      .groups = "drop"
    )
  
  # EIA summary
  eia_summary <- eia_clean %>%
    group_by(tech_group) %>%
    summarise(
      EIA_units = n(),
      EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
      .groups = "drop"
    )
  
  # Join and calculate differences
  full_join(ei_summary, eia_summary, by = "tech_group") %>%
    mutate(
      state = state_name,
      diff_available_capacity = EI_available_capacity - EIA_capacity,
      pct_diff_available_capacity = ifelse(EIA_capacity > 0,
                                           (diff_available_capacity / EIA_capacity) * 100,
                                           NA),
      diff_total_capacity = EI_total_capacity - EIA_capacity,
      pct_diff_total_capacity = ifelse(EIA_capacity > 0,
                                       (diff_total_capacity / EIA_capacity) * 100,
                                       NA)
    ) %>%
    select(state, tech_group, EI_total_units, EI_available_units, EIA_units,
           EI_total_capacity, EI_available_capacity, EIA_capacity,
           diff_total_capacity, pct_diff_total_capacity,
           diff_available_capacity, pct_diff_available_capacity)
}

```

```{r}

ei_va_mapped <- map_technology(VA_ei_gens, "EI")
ei_md_mapped <- map_technology(MD_ei_gens, "EI")

```

### VA 
```{r}
## Run for VA and MD
va_summary_units <- summarize_capacity_units(VA_ei_gens, eia_va, "VA")
va_summary_units
```

### MD

```{r}
md_summary_units <- summarize_capacity_units(MD_ei_gens, eia_md, "MD")

md_summary_units
```

## Capacity Summary Comparison

```{r summary-function}
summarize_capacity_by_state <- function(ei_df, eia_df, state_name) {
  
  # ---- EI: split available vs unavailable ----
  ei_clean <- map_technology(ei_df, "EI")
  
  ei_available <- ei_clean %>%
    filter(available == TRUE)
  
  ei_unavailable <- ei_clean %>%
    filter(available == FALSE)
  
  # ---- Aggregate EI ----
  ei_summary_available <- ei_available %>%
    group_by(tech_group) %>%
    summarise(EI_available_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  ei_summary_total <- ei_clean %>%
    group_by(tech_group) %>%
    summarise(EI_total_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  # ---- EIA ----
  eia_clean <- map_technology(eia_df, "EIA")
  
  eia_summary <- eia_clean %>%
    group_by(tech_group) %>%
    summarise(EIA_capacity = sum(nameplate_capacity, na.rm = TRUE), .groups = "drop")
  
  # ---- Merge ----
  summary <- full_join(ei_summary_available, eia_summary, by = "tech_group") %>%
    full_join(ei_summary_total, by = "tech_group") %>%
    mutate(
      state = state_name,
      EI_available_capacity = replace_na(EI_available_capacity, 0),
      EI_total_capacity = replace_na(EI_total_capacity, 0),
      EIA_capacity = replace_na(EIA_capacity, 0),
      
      # Main comparison (correct one)
      diff_available = EI_available_capacity - EIA_capacity,
      pct_diff_available = ifelse(EIA_capacity > 0,
                                 diff_available / EIA_capacity * 100, NA),
      
      # Diagnostic comparison
      diff_total = EI_total_capacity - EIA_capacity,
      pct_diff_total = ifelse(EIA_capacity > 0,
                             diff_total / EIA_capacity * 100, NA)
    )
  
  return(summary)
}
```


### VA 

```{r run-summary}
va_summary <- summarize_capacity_by_state(VA_ei_gens, eia_va, "VA")
md_summary <- summarize_capacity_by_state(MD_ei_gens, eia_md, "MD")
all_states_summary <- bind_rows(va_summary, md_summary)
va_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "VA Capacity Comparison (Available vs Total EI)")

va_plot_df <- va_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )
ggplot(va_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (VA)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )
```


### MD 
```{r }
md_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "MD Capacity Comparison (Available vs Total EI)")
md_plot_df <- md_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )

ggplot(md_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (MD)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )

```

---

# EI and EIA comparison

```{r}
library(dplyr)
library(readr)


#' Group generators by geographic location and aggregate their capacities
#'
#' @param gens data.frame: Generator dataframe
#' @param lat_col character: Name of latitude column
#' @param lon_col character: Name of longitude column
#' @param capacity_col character: Name of capacity column
#' @param name_col character: Name of generator name column
#'
#' @return data.frame: Grouped generators with aggregated capacity and generator lists
group_generators_by_location = function(gens, lat_col, lon_col, capacity_col, name_col) {
  grouped = gens %>%
    group_by(across(all_of(c(lat_col, lon_col)))) %>%
    summarise(
      gen_names = paste(get(name_col), collapse = "; "),
      n_generators = n(),
      total_capacity = sum(get(capacity_col), na.rm = TRUE),
      first_gen_name = first(get(name_col)),
      .groups = 'drop'
    )
  
  return(grouped)
}

#' Compare generators between EI and EIA datasets using geographic coordinates
#' First groups generators by location, then matches locations
#'
#' @param ei_gens data.frame: EI generators dataframe (must have Lat, Lon, Rating columns)
#' @param eia_gens data.frame: EIA generators dataframe (must have lat, lon, nameplate_capacity columns)
#' @param coord_tolerance numeric: Maximum distance in degrees for matching (default: 0.02 ≈ 2.2km)
#'
#' @return list: Contains grouped and detailed views of matches, plus summary statistics
compare_generator_fleets = function(ei_gens, eia_gens, coord_tolerance = 0.2) {
  
  # Step 1: Group generators by location
  cat("\nGrouping generators by location...\n")
  
  # Group EI generators
  ei_grouped = ei_gens %>%
    group_by(Lat, Lon) %>%
    summarise(
        GenNames         = paste(GenName, collapse = "; "),
        BusNumbers       = paste(BusNumber, collapse = "; "),
        BusNames         = paste(BusName, collapse = "; "),
        n_ei_generators  = n(),
        Rating           = sum(Rating),
        Available        = all(Available),
        Type             = first(Type),
        .groups          = 'drop'
    )
  
  cat("  EI:", nrow(ei_gens), "generators →", nrow(ei_grouped), "unique locations\n")
  
  # Group EIA generators
  eia_grouped = eia_gens %>%
    group_by(lat, lon) %>%
    summarise(
      plant_names = paste(plant, collapse = "; "),
      generator_ids = paste(`generator ID`, collapse = "; "),
      n_eia_generators = n(),
      nameplate_capacity = sum(nameplate_capacity),
      technology = first(technology),
      ba = first(ba),
      .groups = 'drop'
    )
  
  cat("  EIA:", nrow(eia_gens), "generators →", nrow(eia_grouped), "unique locations\n")
  
  # Step 2: Match locations
  cat("\nMatching locations (tolerance:", coord_tolerance, "°)...\n")
  
  # Function to find closest match
  find_closest_match = function(lat, lon, df, tolerance) {
    distances = sqrt((df$lat - lat)^2 + (df$lon - lon)^2)
    min_dist = min(distances)
    idx = which.min(distances)
    
    if (min_dist <= tolerance) {
      return(idx)
    } else {
      return(NA)
    }
  }
  
  # Match EI locations to EIA locations
  ei_grouped$eia_match_idx = sapply(1:nrow(ei_grouped), function(i) {
    find_closest_match(ei_grouped$Lat[i], ei_grouped$Lon[i], 
                      eia_grouped, coord_tolerance)
  })
  
  # Step 3: Create matched dataset
  in_both_grouped = ei_grouped %>% 
    filter(!is.na(eia_match_idx))
  
  if (nrow(in_both_grouped) > 0) {
    # Add EIA data
    in_both_grouped$eia_lat = eia_grouped$lat[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_lon = eia_grouped$lon[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_plant_names = eia_grouped$plant_names[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_generator_ids = eia_grouped$generator_ids[in_both_grouped$eia_match_idx]
    in_both_grouped$n_eia_generators = eia_grouped$n_eia_generators[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_capacity = eia_grouped$nameplate_capacity[in_both_grouped$eia_match_idx]
    
    # Calculate differences
    in_both_grouped$capacity_diff = in_both_grouped$Rating - in_both_grouped$eia_capacity
    in_both_grouped$capacity_abs_diff = abs(in_both_grouped$capacity_diff)
    
    # Percentage difference with respect to EIA capacity
    in_both_grouped$capacity_pct_diff = (in_both_grouped$Rating - in_both_grouped$eia_capacity) / 
                                         in_both_grouped$eia_capacity * 100
    in_both_grouped$capacity_abs_pct_diff = abs(in_both_grouped$capacity_pct_diff)
    in_both_grouped$capacity_ratio = in_both_grouped$Rating / in_both_grouped$eia_capacity
    
    
    
    # Rename for clarity
    in_both_grouped = in_both_grouped %>%
      rename(ei_lat = Lat, ei_lon = Lon)
  }
  
  # Step 4: Create detailed view (expand to individual generators)
  in_both_detailed = data.frame()
  
  if (nrow(in_both_grouped) > 0) {
    for (i in 1:nrow(in_both_grouped)) {
      row = in_both_grouped[i, ]
      
      # Get individual EI generators at this location
      ei_at_location = ei_gens %>%
        filter(Lat == row$ei_lat & Lon == row$ei_lon)
      
      # Add aggregated matching info to each
      ei_at_location$eia_plant_names = row$eia_plant_names
      ei_at_location$eia_generator_ids = row$eia_generator_ids
      ei_at_location$n_eia_generators = row$n_eia_generators
      ei_at_location$n_ei_at_location = row$n_ei_generators
      ei_at_location$eia_capacity = row$eia_capacity
      ei_at_location$total_ei_capacity_at_location = row$Rating
      ei_at_location$eia_lat = row$eia_lat
      ei_at_location$eia_lon = row$eia_lon
      ei_at_location$capacity_diff = row$capacity_diff
      ei_at_location$capacity_abs_diff = row$capacity_abs_diff
      
      # Rename lat/lon
      ei_at_location = ei_at_location %>%
        rename(ei_lat = Lat, ei_lon = Lon)
      
      in_both_detailed = bind_rows(in_both_detailed, ei_at_location)
    }
  }
  
  # Step 5: Identify unmatched generators
  in_ei_only_grouped = ei_grouped %>% 
    filter(is.na(eia_match_idx))
  
  matched_eia_indices = in_both_grouped$eia_match_idx[!is.na(in_both_grouped$eia_match_idx)]
  in_eia_only_grouped = eia_grouped %>% 
    filter(!(row_number() %in% matched_eia_indices))
  
  # Calculate summary statistics
  total_ei_capacity = sum(ei_gens$Rating, na.rm = TRUE)
  total_eia_capacity = sum(eia_gens$nameplate_capacity, na.rm = TRUE)
  matched_ei_capacity = if(nrow(in_both_grouped) > 0) sum(in_both_grouped$Rating, na.rm = TRUE) else 0
  matched_eia_capacity = if(nrow(in_both_grouped) > 0) sum(in_both_grouped$eia_capacity, na.rm = TRUE) else 0
  mean_capacity_diff = if(nrow(in_both_grouped) > 0) mean(in_both_grouped$capacity_diff, na.rm = TRUE) else 0
  mean_capacity_abs_diff = if(nrow(in_both_grouped) > 0) mean(in_both_grouped$capacity_abs_diff, na.rm = TRUE) else 0
  
  # Print summary
  cat("\n", paste(rep("=", 60), collapse = ""), "\n")
  cat("Fleet Comparison Results\n")
  cat(paste(rep("=", 60), collapse = ""), "\n")
  cat("Matched locations:", nrow(in_both_grouped), "\n")
  if(nrow(in_both_grouped) > 0) {
    cat("  Total EI generators at matched locations:", sum(in_both_grouped$n_ei_generators), "\n")
    cat("  Total EIA generators at matched locations:", sum(in_both_grouped$n_eia_generators), "\n")
  }
  cat("Unmatched EI locations:", nrow(in_ei_only_grouped), "\n")
  cat("Unmatched EIA locations:", nrow(in_eia_only_grouped), "\n")
  cat("\nCapacity Summary:\n")
  cat("  Total EI capacity:", round(total_ei_capacity, 2), "MW\n")
  cat("  Total EIA capacity:", round(total_eia_capacity, 2), "MW\n")
  cat("  Matched EI capacity:", round(matched_ei_capacity, 2), "MW\n")
  cat("  Matched EIA capacity:", round(matched_eia_capacity, 2), "MW\n")
  cat("  Mean capacity difference:", round(mean_capacity_diff, 2), "MW\n")
  cat("  Mean absolute capacity difference:", round(mean_capacity_abs_diff, 2), "MW\n")
  cat(paste(rep("=", 60), collapse = ""), "\n\n")
  
  # Return results as a list
  return(list(
    in_both_grouped = in_both_grouped,
    in_both_detailed = in_both_detailed,
    in_ei_only_grouped = in_ei_only_grouped,
    in_eia_only_grouped = in_eia_only_grouped,
    ei_grouped = ei_grouped,
    eia_grouped = eia_grouped,
    summary = list(
      total_ei_capacity = total_ei_capacity,
      total_eia_capacity = total_eia_capacity,
      matched_ei_capacity = matched_ei_capacity,
      matched_eia_capacity = matched_eia_capacity,
      mean_capacity_diff = mean_capacity_diff,
      mean_capacity_abs_diff = mean_capacity_abs_diff,
      n_matched_locations = nrow(in_both_grouped),
      n_ei_only_locations = nrow(in_ei_only_grouped),
      n_eia_only_locations = nrow(in_eia_only_grouped)
    )
  ))
}


```

```{r}
#' Compare generator fleets for a specific technology type
#'
#' @param ei_gens data.frame: Full EI generators dataframe (virginia_resources.csv)
#' @param eia_gens data.frame: Full EIA generators dataframe
#' @param type character: "Solar", "Wind", "Hydro", "Thermal", "Storage"
#' @param coord_tolerance numeric: tolerance in degrees (default 0.2)
compare_by_type = function(ei_gens, eia_gens, type = "Solar", coord_tolerance = 0.2,
                           eia_type_col = "technology") {
  
  # ── 1. Type → filter mappings ──────────────────────────────────────────────
  ei_type_map = list(
  "Solar"   = c("Solar"),
  "Wind"    = c("Wind"),
  "Hydro"   = c("Hydro"),
  "Thermal" = c("ThermalStandard", "ThermalMultiStart"),
  "Nuclear" = c("ThermalStandard"),   # 👈 special handling
  "Storage" = c("Storage")
)
  
  if (type == "Nuclear") {
  ei_filtered = ei_gens %>%
    filter(generator_type == "ThermalStandard", fuel == "NUCLEAR")
}
  
  eia_type_map = list(
  "Solar"   = c("Solar Photovoltaic"),
  "Wind"    = c("Onshore Wind Turbine", "Offshore Wind Turbine"),
  "Hydro"   = c("Conventional Hydroelectric", "Run-of-River Hydroelectric", "Hydro"),
  "Thermal" = c("Natural Gas Steam Turbine", "Natural Gas Fired Combined Cycle",
                "Natural Gas Fired Combustion Turbine", "Coal Steam Turbine"),
  "Nuclear" = c("Nuclear"),
  "Storage" = c("Batteries")
)
  
  if (!type %in% names(ei_type_map)) {
    stop("Invalid type. Choose from: ", paste(names(ei_type_map), collapse = ", "))
  }
  
  # ── 2. Filter by type ──────────────────────────────────────────────────────
  ei_filtered = ei_gens %>%
  filter(
    generator_type %in% ei_type_map[[type]] &
    !(type == "Thermal" & fuel == "NUCLEAR")
  )
  
  # Use column override if needed
  if (!eia_type_col %in% colnames(eia_gens)) {
    stop(paste("EIA type column", eia_type_col, "not found in EIA dataset"))
  }
  
  eia_filtered = eia_gens %>% filter(.data[[eia_type_col]] %in% eia_type_map[[type]])
  
  cat("=== Type:", type, "===\n")
  cat("EI generators:  ", nrow(ei_filtered), "\n")
  cat("EIA generators: ", nrow(eia_filtered), "\n")
  
  if (nrow(ei_filtered) == 0 || nrow(eia_filtered) == 0) {
    warning("No generators found for type: ", type)
    return(NULL)
  }
  
  # ── 3. Rename EI columns to match compare_generator_fleets expectations ────
  ei_renamed = ei_filtered %>%
    rename(
      GenName   = name,
      BusNumber = bus_number,
      BusName   = bus,
      Lat       = lat,
      Lon       = lon,
      Rating    = rating,
      Available = available,
      Type      = generator_type
    )
  
  # ── 4. Run comparison ──────────────────────────────────────────────────────
  result = compare_generator_fleets(
    ei_gens        = ei_renamed,
    eia_gens       = eia_filtered,
    coord_tolerance = coord_tolerance
  )
  
  return(result)
}
```

## VA

### Solar

```{r}
solar_results = compare_by_type(VA_ei_gens, eia_va_renew, type = "Solar", coord_tolerance = 0.2)


# Access results
solar_results$in_both_grouped%>%
  dplyr::select(n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
                Rating, eia_capacity, 
                capacity_diff, capacity_abs_diff,
                capacity_pct_diff, capacity_abs_pct_diff,   
                ei_lat, eia_lat, ei_lon, eia_lon) %>%
  dplyr::arrange((capacity_abs_pct_diff))
```

### Wind

```{r}
wind_results  = compare_by_type(VA_ei_gens, eia_va_renew, type = "Wind", coord_tolerance = 0.2)

# wind_results$in_both_grouped %>%
#   dplyr::select(n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
#                 Rating, eia_capacity, 
#                 capacity_diff, capacity_abs_diff,
#                 capacity_pct_diff, capacity_abs_pct_diff,   
#                 ei_lat, eia_lat, ei_lon, eia_lon) %>%
#   dplyr::arrange(desc(capacity_abs_pct_diff))
```

### Hydro

```{r}

eia_va_hydro_tech <- eia_va_hydro %>%
  mutate(technology = "Hydro")
hydro_results = compare_by_type(VA_ei_gens, eia_va_hydro_tech, type = "Hydro", coord_tolerance = 0.2)

# Check for matches
if (!is.null(hydro_results) && !is.null(hydro_results$in_both_grouped)) {
  hydro_results$in_both_grouped %>%
    dplyr::select(
      n_eia_generators, eia_plant_names, 
      n_ei_generators, BusNames,
      Rating, eia_capacity, 
      capacity_diff, capacity_abs_diff,
      capacity_pct_diff, capacity_abs_pct_diff,   
      ei_lat, eia_lat, ei_lon, eia_lon
    ) %>%
    dplyr::arrange((capacity_abs_pct_diff))
} else {
  cat("No matched Hydro generators found.\n")
}
```

### Thermal

```{r}
# eia_va_thermal_tech <- eia_va_thermal %>%
#   mutate(technology = "Thermal")
# thermal_results = compare_by_type(
#   VA_ei_gens,
#   eia_va_thermal_tech,
#   type = "Thermal",
#   coord_tolerance = 0.2,
#   eia_type_col = "technology",
#   eia_type_values = c(
#     "Natural Gas Steam Turbine",
#     "Natural Gas Fired Combined Cycle",
#     "Natural Gas Fired Combustion Turbine",
#     "Coal Steam Turbine"
#   )
# )
# 
# thermal_results$in_both_grouped %>%
#   dplyr::select(
#     n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
#     Rating, eia_capacity, 
#     capacity_diff, capacity_abs_diff,
#     capacity_pct_diff, capacity_abs_pct_diff,   
#     ei_lat, eia_lat, ei_lon, eia_lon
#   ) %>%
#   dplyr::arrange((capacity_abs_pct_diff))
```

