---
title: "Data Cleaning & Alignment (EI/Sienna ↔ EIA)"
author: "Sara Abril-Guevara"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: true
    toc_float: true
    df_print: paged
    number_sections: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) 
library(readr) 
library(ggplot2) 
library(tidyr) 
library(knitr)
library(readxl)
library(stringr)
library(dplyr)
library(readr)
library(stringdist)
library(purrr)
```
# Overview

This document summarizes the data cleaning and validation process used to align generator fleets from EI/Sienna and EIA datasets for selected states (MN, IL, MD, VA, WV).

The goal is to ensure consistency in installed capacity before running simulation models.


## Methodology

- DONE: Extract EI/Sienna generators and bus attributes using Kestrel

- DONE: Aggregate generators by geographic coordinates

- DONE: Standardize technology categories across datasets

- WIP: Compare installed capacity by state and technology

- WIP: Flag differences for further spatial matching and manual review

```{r cars, message = FALSE}
# EI datasets 
VA_ei_gens = read_csv("~/Downloads/virginia_resources.csv") 
MD_ei_gens = read_csv("~/Downloads/maryland_resources.csv")
WV_ei_gens = read_csv("~/Downloads/west_virginia_resources.csv")

# EI updated datasets 
VA_updated_ei_gens = read_csv("/Users/sabrilg/Documents/GitHub/va_updates/virginia_updated_resources.csv") 
MD_updated_ei_gens = read_csv("/Users/sabrilg/Documents/GitHub/va_updates/maryland_updated_resources.csv")
WV_updated_ei_gens = read_csv("/Users/sabrilg/Documents/GitHub/va_updates/west_virginia_updated_resources.csv")

#EIA
eia_all <- read_excel("3_1_Generator_Y2024.xlsx", sheet = "Operable", skip = 1, col_types = "text")

eia_all <- eia_all %>%
  mutate(
    `Nameplate Capacity (MW)` = as.numeric(`Nameplate Capacity (MW)`),
    `Summer Capacity (MW)` = as.numeric(`Summer Capacity (MW)`),
    `Winter Capacity (MW)` = as.numeric(`Winter Capacity (MW)`),
    `Minimum Load (MW)` = as.numeric(`Minimum Load (MW)`),
    `Nameplate Power Factor` = as.numeric(`Nameplate Power Factor`)
  ) %>%
  rename(nameplate_capacity = `Nameplate Capacity (MW)`)
eia_va <- eia_all %>% filter(State == "VA")
eia_md <- eia_all %>% filter(State == "MD")
eia_wv <- eia_all %>% filter(State == "WV")

```



```{r}
# Function to map EIA generators to high-level tech groups
map_tech_group <- function(df) {
  df %>%
    mutate(
      tech_group = case_when(
        # Nuclear
        `Prime Mover` == "NU" | Technology == "Nuclear" ~ "Nuclear",
        
        # Pumped Storage
        `Prime Mover` == "PS" | str_detect(Technology, "Pumped Storage") ~ "Pumped Storage",
        
        # Hydro (excluding Pumped Storage)
        `Prime Mover` == "HY" | str_detect(Technology, "Hydroelectric") & !str_detect(Technology, "Pumped Storage") ~ "Hydro",
        
        # Wind
        `Prime Mover` == "WT" | str_detect(Technology, "Wind") ~ "Wind",
        
        # Solar
        `Prime Mover` %in% c("PV", "CE") | str_detect(Technology, "Solar") ~ "Solar",
        
        # Thermal / Fossil
        `Prime Mover` %in% c("ST", "CT", "CA", "GT", "IC", "CS", "FC", "CP") |
          Technology %in% c(
            "Conventional Steam Coal", 
            "Natural Gas Fired Combined Cycle", 
            "Natural Gas Steam Turbine",
            "Coal Integrated Gasification Combined Cycle",
            "Natural Gas Fired Combustion Turbine",
            "Natural Gas Internal Combustion Engine"
          ) ~ "Thermal",
        
        # Batteries / Storage / Flywheels
        `Prime Mover` %in% c("BA", "BT", "FW") | str_detect(Technology, "Battery|Flywheels|Storage") ~ "Batteries",
        
        # Biomass / Waste / Geothermal / Other
        str_detect(Technology, "Biomass|Waste|Geothermal|Other") ~ "Other",
        
        # Fallback
        TRUE ~ "Other"
      )
    )
}
# Clean and map tech_group for all three states
eia_va <- eia_va %>% map_tech_group()
eia_md <- eia_md %>% map_tech_group()
eia_wv <- eia_wv %>% map_tech_group()
```




## Technology Mapping

```{r}
map_technology = function(df, source = "EI") {
  
  if (source == "EI") {
    
    df %>%
      mutate(
        tech_group = case_when(
          fuel == "NUCLEAR" ~ "Nuclear",
          generator_type %in% c("Solar") ~ "Solar",
          generator_type %in% c("Wind") ~ "Wind",
          generator_type %in% c("ThermalStandard", "ThermalMultiStart") ~ "Thermal",
          generator_type %in% c("Storage") ~ "Storage",
          prime_mover_type == "PS" ~ "Pumped Storage",
          generator_type %in% c("Hydro") ~ "Hydro",
          TRUE ~ "Other"
        )
      )
    
  } else {
    
    df %>%
      mutate(
        tech_group = case_when(
          
          # --- Explicit first (avoid misclassification) ---
          Technology == "Nuclear" ~ "Nuclear",
          
          Technology == "Hydroelectric Pumped Storage" | `Prime Mover` == "PS" ~ "Pumped Storage",
          
          Technology == "Conventional Hydroelectric" ~ "Hydro",
          
          Technology == "Batteries" ~ "Batteries",
          
          Technology %in% c("Flywheels", "Natural Gas with Compressed Air Storage") ~ "Storage",
          
          # --- Pattern-based (your original strength) ---
          grepl("Solar", Technology) ~ "Solar",
          grepl("Wind", Technology) ~ "Wind",
          
          grepl("Hydro", Technology) & !grepl("Pumped Storage", Technology) ~ "Hydro",
          
          grepl("Gas|Coal|Petroleum|Steam|Combustion", Technology) ~ "Thermal",
          
          grepl("Biomass|Waste|Landfill|Geothermal", Technology) ~ "Other",
          
          TRUE ~ "Other"
        )
      )
  }
}
# Summarize function including units for both EI and EIA
summarize_capacity_units = function(ei_df, eia_df, state_name) {

  ei_clean  = map_technology(ei_df, "EI")
  eia_clean = map_technology(eia_df, "EIA")

  # EI summary
  ei_summary = ei_clean %>%
    group_by(tech_group) %>%
    summarise(
      EI_total_units = n(),
      EI_available_units = sum(available == TRUE, na.rm = TRUE),
      EI_total_capacity = sum(rating, na.rm = TRUE),
      EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      .groups = "drop"
    )

  # EIA summary
  eia_summary = eia_clean %>%
    group_by(tech_group) %>%
    summarise(
      EIA_units = n(),
      EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
      .groups = "drop"
    )

  # Join summaries and calculate differences
  summary_df = full_join(ei_summary, eia_summary, by = "tech_group") %>%
    mutate(
      state = state_name,
      diff_available_capacity = EI_available_capacity - EIA_capacity,
      pct_diff_available_capacity = ifelse(EIA_capacity > 0,
                                           (diff_available_capacity / EIA_capacity) * 100,
                                           NA),
      diff_total_capacity = EI_total_capacity - EIA_capacity,
      pct_diff_total_capacity = ifelse(EIA_capacity > 0,
                                       (diff_total_capacity / EIA_capacity) * 100,
                                       NA)
    ) %>%
    select(state, tech_group, EI_total_units, EI_available_units, EIA_units,
           EI_total_capacity, EI_available_capacity, EIA_capacity,
           diff_total_capacity, pct_diff_total_capacity,
           diff_available_capacity, pct_diff_available_capacity)

  # Optional: reorder tech groups
  tech_order = c("Nuclear", "Thermal", "Hydro", "Pumped Storage", "Storage", "Solar", "Wind", "Other")
  summary_df$tech_group = factor(summary_df$tech_group, levels = tech_order)
  summary_df = summary_df %>% arrange(tech_group)

  return(summary_df)
}
```



### VA 
```{r}
## Run for VA
va_summary_units = summarize_capacity_units(VA_ei_gens, eia_va, "VA")
va_summary_units
va_up_summary_units = summarize_capacity_units(VA_updated_ei_gens, eia_va, "VA")
va_up_summary_units
```


### MD

```{r}
md_summary_units = summarize_capacity_units(MD_ei_gens, eia_md, "MD")
md_summary_units
md_up_summary_units = summarize_capacity_units(MD_updated_ei_gens, eia_md, "MD")
md_up_summary_units
```

### WV

```{r}
wv_summary_units = summarize_capacity_units(WV_ei_gens, eia_wv, "WV")
wv_summary_units
wv_up_summary_units = summarize_capacity_units(WV_updated_ei_gens, eia_wv, "WV")
wv_up_summary_units
```

## Capacity Summary Comparison

```{r summary-function}
summarize_capacity_by_state = function(ei_df, eia_df, state_name) {
  
  # ---- EI: split available vs unavailable ----
  ei_clean = map_technology(ei_df, "EI")
  
  ei_available = ei_clean %>%
    filter(available == TRUE)
  
  ei_unavailable = ei_clean %>%
    filter(available == FALSE)
  
  # ---- Aggregate EI ----
  ei_summary_available = ei_available %>%
    group_by(tech_group) %>%
    summarise(EI_available_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  ei_summary_total = ei_clean %>%
    group_by(tech_group) %>%
    summarise(EI_total_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  # ---- EIA ----
  eia_clean = map_technology(eia_df, "EIA")
  
  eia_summary = eia_clean %>%
    group_by(tech_group) %>%
    summarise(EIA_capacity = sum(nameplate_capacity, na.rm = TRUE), .groups = "drop")
  
  # ---- Merge ----
  summary = full_join(ei_summary_available, eia_summary, by = "tech_group") %>%
    full_join(ei_summary_total, by = "tech_group") %>%
    mutate(
      state = state_name,
      EI_available_capacity = replace_na(EI_available_capacity, 0),
      EI_total_capacity = replace_na(EI_total_capacity, 0),
      EIA_capacity = replace_na(EIA_capacity, 0),
      
      # Main comparison (correct one)
      diff_available = EI_available_capacity - EIA_capacity,
      pct_diff_available = ifelse(EIA_capacity > 0,
                                 diff_available / EIA_capacity * 100, NA),
      
      # Diagnostic comparison
      diff_total = EI_total_capacity - EIA_capacity,
      pct_diff_total = ifelse(EIA_capacity > 0,
                             diff_total / EIA_capacity * 100, NA)
    )
  
  return(summary)
}
```

```{r}
va_summary = summarize_capacity_by_state(VA_ei_gens, eia_va, "VA")
md_summary = summarize_capacity_by_state(MD_ei_gens, eia_md, "MD")
wv_summary = summarize_capacity_by_state(WV_ei_gens, eia_wv, "WV")
all_states_summary = bind_rows(va_summary, md_summary, wv_summary)

# UPDATED summaries
va_updated_summary = summarize_capacity_by_state(VA_updated_ei_gens, eia_va, "VA_updated")
md_updated_summary = summarize_capacity_by_state(MD_updated_ei_gens, eia_md, "MD_updated")
wv_updated_summary = summarize_capacity_by_state(WV_updated_ei_gens, eia_wv, "WV_updated")
```

```{r}
plot_thermal_discrepancy <- function(ei_df, eia_df, state_name, available_only = TRUE) {
  
  # -------------------------
  # EI Thermal (EXCLUDE NUCLEAR)
  # -------------------------
  ei_thermal <- ei_df %>%
    filter(prime_mover_type %in% c("ST", "CT", "CC", "OT")) %>%
    filter(fuel != "NUCLEAR") %>%   # <-- key fix
    { if (available_only) filter(., available == TRUE) else . } %>%
    group_by(fuel) %>%
    summarise(EI_capacity = sum(rating, na.rm = TRUE), .groups = "drop")
  
  # -------------------------
  # EIA Thermal
  # -------------------------
  eia_thermal <- eia_df %>%
    filter(Technology %in% c(
      "Conventional Steam Coal",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Petroleum Liquids",
      "Petroleum Coke",
      "Other Gases"
    )) %>%
    mutate(fuel = case_when(
      Technology == "Conventional Steam Coal" ~ "COAL",
      
      Technology %in% c(
        "Natural Gas Fired Combined Cycle",
        "Natural Gas Steam Turbine",
        "Natural Gas Fired Combustion Turbine",
        "Natural Gas Internal Combustion Engine",
        "Other Gases"
      ) ~ "NATURAL_GAS",
      
      Technology %in% c(
        "Petroleum Liquids",
        "Petroleum Coke"
      ) ~ "DISTILLATE_FUEL_OIL",
      
      TRUE ~ "OTHER"
    )) %>%
    group_by(fuel) %>%
    summarise(EIA_capacity = sum(nameplate_capacity, na.rm = TRUE), .groups = "drop")
  
  # -------------------------
  # Combine
  # -------------------------
  thermal_compare <- full_join(ei_thermal, eia_thermal, by = "fuel") %>%
    replace_na(list(EI_capacity = 0, EIA_capacity = 0)) %>%
    pivot_longer(cols = c(EI_capacity, EIA_capacity),
                 names_to = "source",
                 values_to = "capacity")
  
  # -------------------------
  # Plot
  # -------------------------
  ggplot(thermal_compare, aes(x = fuel, y = capacity, fill = source)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    theme_minimal() +
    labs(
      title = paste(state_name, "- Thermal Capacity by Fuel (EI (Available) vs EIA)"),
      x = "Fuel Type",
      y = "Capacity (MW)",
      fill = "Source"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 9)
    ) +
    scale_fill_manual(values = c(
      "EI_capacity" = "#FF7F73",
      "EIA_capacity" = "#00C4CC"
    ))
}


```

```{r}
plot_thermal_3bars <- function(ei_df, eia_df, state_name) {
  
  # -------------------------
  # EI Thermal
  # -------------------------
  ei_thermal <- ei_df %>%
    filter(prime_mover_type %in% c("ST", "CT", "CC", "OT")) %>%
    filter(fuel != "NUCLEAR") %>%
    group_by(fuel) %>%
    summarise(
      EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      EI_total_capacity     = sum(rating, na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # EIA Thermal
  # -------------------------
  eia_thermal <- eia_df %>%
    filter(Technology %in% c(
      "Conventional Steam Coal",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Petroleum Liquids",
      "Petroleum Coke",
      "Other Gases"
    )) %>%
    mutate(fuel = case_when(
      Technology == "Conventional Steam Coal" ~ "COAL",
      
      Technology %in% c(
        "Natural Gas Fired Combined Cycle",
        "Natural Gas Steam Turbine",
        "Natural Gas Fired Combustion Turbine",
        "Natural Gas Internal Combustion Engine",
        "Other Gases"
      ) ~ "NATURAL_GAS",
      
      Technology %in% c(
        "Petroleum Liquids",
        "Petroleum Coke"
      ) ~ "DISTILLATE_FUEL_OIL",
      
      TRUE ~ "OTHER"
    )) %>%
    group_by(fuel) %>%
    summarise(
      EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # Combine
  # -------------------------
  thermal_compare <- full_join(ei_thermal, eia_thermal, by = "fuel") %>%
    replace_na(list(
      EI_available_capacity = 0,
      EI_total_capacity = 0,
      EIA_capacity = 0
    )) %>%
    pivot_longer(
      cols = c(EI_available_capacity, EI_total_capacity, EIA_capacity),
      names_to = "source",
      values_to = "capacity"
    )
  
  # -------------------------
  # Plot
  # -------------------------
  ggplot(thermal_compare, aes(x = fuel, y = capacity, fill = source)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    theme_minimal() +
    labs(
      title = paste(state_name, "- Thermal Capacity (EI Available vs Total vs EIA)"),
      x = "Fuel Type",
      y = "Capacity (MW)",
      fill = "Source"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 9)
    ) +
    scale_fill_manual(values = c(
      "EI_available_capacity" = "#FF7F73",   # coral
      "EI_total_capacity"     = "#FFA07A",   # lighter coral
      "EIA_capacity"          = "#00C4CC"    # teal
    ))
}
```

```{r}
plot_compare_versions <- function(original_df, updated_df, state_name) {
  
  original_long <- original_df %>%
    select(tech_group, EI_available_capacity, EIA_capacity) %>%
    mutate(version = "Original EI")
  
  updated_long <- updated_df %>%
    select(tech_group, EI_available_capacity, EIA_capacity) %>%
    mutate(version = "Updated EI")
  
  combined <- bind_rows(original_long, updated_long) %>%
    pivot_longer(
      cols = c(EI_available_capacity, EIA_capacity),
      names_to = "source",
      values_to = "capacity"
    ) %>%
    mutate(
  source = case_when(
    source == "EI_available_capacity" & version == "Original EI" ~ "EI avail (Original)",
    source == "EI_available_capacity" & version == "Updated EI" ~ "EI avail (Updated)",
    source == "EIA_capacity" ~ "EIA"
  )
)
  
  ggplot(combined, aes(x = tech_group, y = capacity, fill = source)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    theme_minimal() +
    labs(
      title = paste(state_name, "- Capacity Comparison (Original vs Updated EI vs EIA)"),
      x = "Technology",
      y = "Capacity (MW)",
      fill = "Source"
    ) +
    scale_fill_manual(values = c(
      "EI avail (Original)" = "#FF7F73",
      "EI avail (Updated)"  = "#FF4FF3",
      "EIA"           = "#00C4CC"
    )) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r}
plot_thermal_4bars <- function(ei_df, ei_updated_df, eia_df, state_name) {
  
  # -------------------------
  # EI Thermal (ORIGINAL)
  # -------------------------
  ei_thermal <- ei_df %>%
    filter(prime_mover_type %in% c("ST", "CT", "CC", "OT")) %>%
    filter(fuel != "NUCLEAR") %>%
    group_by(fuel) %>%
    summarise(
      EI_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      EI_total_capacity     = sum(rating, na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # EI Thermal (UPDATED availability)
  # -------------------------
  ei_updated_thermal <- ei_updated_df %>%
    filter(prime_mover_type %in% c("ST", "CT", "CC", "OT")) %>%
    filter(fuel != "NUCLEAR") %>%
    group_by(fuel) %>%
    summarise(
      EI_updated_available_capacity = sum(rating[available == TRUE], na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # EIA Thermal
  # -------------------------
  eia_thermal <- eia_df %>%
    filter(Technology %in% c(
      "Conventional Steam Coal",
      "Natural Gas Fired Combined Cycle",
      "Natural Gas Steam Turbine",
      "Natural Gas Fired Combustion Turbine",
      "Natural Gas Internal Combustion Engine",
      "Petroleum Liquids",
      "Petroleum Coke",
      "Other Gases"
    )) %>%
    mutate(fuel = case_when(
      Technology == "Conventional Steam Coal" ~ "COAL",
      
      Technology %in% c(
        "Natural Gas Fired Combined Cycle",
        "Natural Gas Steam Turbine",
        "Natural Gas Fired Combustion Turbine",
        "Natural Gas Internal Combustion Engine",
        "Other Gases"
      ) ~ "NATURAL_GAS",
      
      Technology %in% c(
        "Petroleum Liquids",
        "Petroleum Coke"
      ) ~ "DISTILLATE_FUEL_OIL",
      
      TRUE ~ "OTHER"
    )) %>%
    group_by(fuel) %>%
    summarise(
      EIA_capacity = sum(nameplate_capacity, na.rm = TRUE),
      .groups = "drop"
    )
  
  # -------------------------
  # Combine all
  # -------------------------
  thermal_compare <- ei_thermal %>%
    full_join(ei_updated_thermal, by = "fuel") %>%
    full_join(eia_thermal, by = "fuel") %>%
    replace_na(list(
      EI_available_capacity = 0,
      EI_total_capacity = 0,
      EI_updated_available_capacity = 0,
      EIA_capacity = 0
    )) %>%
    pivot_longer(
      cols = c(
        EI_available_capacity,
        EI_updated_available_capacity,
        EI_total_capacity,
        EIA_capacity
      ),
      names_to = "source",
      values_to = "capacity"
    )
  
  # -------------------------
  # Plot
  # -------------------------
  ggplot(thermal_compare, aes(x = fuel, y = capacity, fill = source)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    theme_minimal() +
    labs(
      title = paste(state_name, "- Thermal Capacity (EI Original vs Updated vs Total vs EIA)"),
      x = "Fuel Type",
      y = "Capacity (MW)",
      fill = "Source"
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 9)
    ) +
    scale_fill_manual(values = c(
      "EI_available_capacity"          =  "#FF7F73",  # light coral (original)
      "EI_updated_available_capacity"  = "#FF4FF3",  # darker coral (updated)
      "EI_total_capacity"              = "#FFA07A",  # dark red
      "EIA_capacity"                   = "#00C4CC"   # teal
    ))
   
}
```



### VA 

```{r run-summary}

va_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "VA Capacity Comparison (Available vs Total EI)")

va_plot_df = va_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )
ggplot(va_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (VA)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )
plot_compare_versions(va_summary, va_updated_summary, "Virginia")
```

```{r}
plot_thermal_discrepancy(VA_ei_gens, eia_va, "Virginia")
```


```{r}
plot_thermal_3bars(VA_ei_gens, eia_va, "VA")
```

```{r}
plot_thermal_4bars(VA_ei_gens, VA_updated_ei_gens, eia_va, "VA")
```


### MD 
```{r }
md_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "MD Capacity Comparison (Available vs Total EI)")
md_plot_df = md_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )

ggplot(md_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (MD)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )
plot_compare_versions(md_summary, md_updated_summary, "Maryland")
```

```{r}
plot_thermal_discrepancy(MD_ei_gens, eia_md, "Maryland")
```

```{r}
plot_thermal_3bars(MD_ei_gens, eia_md, "MD")
```

```{r}
plot_thermal_4bars(MD_ei_gens, MD_updated_ei_gens, eia_md, "MD")
```



### WV 
```{r }
wv_summary %>%
  arrange(state, tech_group) %>%
  dplyr::select(
    state, tech_group,
    EI_available_capacity,
    EIA_capacity,
    pct_diff_available,
    EI_total_capacity,
    pct_diff_total
  ) %>%
  kable(digits = 1, caption = "WV Capacity Comparison (Available vs Total EI)")
wv_plot_df = wv_summary %>%
  select(state, tech_group, EI_available_capacity, EIA_capacity) %>%
  pivot_longer(
    cols = c(EI_available_capacity, EIA_capacity),
    names_to = "source",
    values_to = "capacity"
  )

ggplot(wv_plot_df, aes(x = tech_group, y = capacity, fill = source)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  theme_minimal() +
  labs(
    title = "Installed Capacity Comparison (WV)",
    y = "MW",
    x = "Technology",
    fill = "Source"
  )
plot_compare_versions(wv_summary, wv_updated_summary, "West Virginia")
```

```{r}
plot_thermal_discrepancy(WV_ei_gens, eia_wv, "West Virginia")
```

```{r}
plot_thermal_3bars(WV_ei_gens, eia_wv, "WV")
```

```{r}
plot_thermal_4bars(WV_ei_gens, WV_updated_ei_gens, eia_wv, "WV")
```

