---
title: "VA Generator Fleet Comparison: EI vs EIA"
author: "Sara Abril-Guevara"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: true
    toc_float: true
    df_print: paged
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document compares generator fleets between the Energy Information Administration (EIA) and Energy Exemplar (EI) datasets. The comparison is based on geographic coordinates (latitude and longitude) with a configurable tolerance.

## Load Libraries

```{r load-libraries, include=FALSE}
library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
```

## Define Comparison Function

```{r define-function, results='hide'}
#' Compare generators between EI and EIA datasets using geographic coordinates
#'
#' @param ei_gens data.frame: EI generators dataframe (must have Lat, Lon, Rating columns)
#' @param eia_gens data.frame: EIA generators dataframe (must have lat, lon, nameplate_capacity columns)
#' @param coord_tolerance numeric: Maximum distance in degrees for matching (default: 0.02 ≈ 2.2km)
#'
#' @return list: Contains in_both, in_ei_only, in_eia_only data frames and summary statistics
compare_generator_fleets = function(ei_gens, eia_gens, coord_tolerance = 0.02) {
  
  # Function to find matches based on lat/lon
  find_closest_match = function(lat, lon, df, tolerance) {
    distances = sqrt((df$lat - lat)^2 + (df$lon - lon)^2)
    min_dist = min(distances)
    idx = which.min(distances)
    
    if (min_dist <= tolerance) {
      return(idx)
    } else {
      return(NA)
    }
  }
  
  # Create a copy to avoid modifying original
  ei_gens_copy = ei_gens
  
  # Create matching results
  ei_gens_copy$eia_match_idx = sapply(1:nrow(ei_gens_copy), function(i) {
    find_closest_match(ei_gens_copy$Lat[i], ei_gens_copy$Lon[i], 
                      eia_gens, coord_tolerance)
  })
  
  # Generators in both datasets
  in_both = ei_gens_copy %>% filter(!is.na(eia_match_idx))
  
  # Add EIA data to matched generators
  if (nrow(in_both) > 0) {
    in_both$eia_lat = eia_gens$lat[in_both$eia_match_idx]
    in_both$eia_lon = eia_gens$lon[in_both$eia_match_idx]
    in_both$eia_capacity = eia_gens$nameplate_capacity[in_both$eia_match_idx]
    in_both$capacity_diff = in_both$Rating - in_both$eia_capacity
    
    # Rename EI lat/lon for clarity
    in_both = in_both %>%
      rename(ei_lat = Lat, ei_lon = Lon)
  }
  
  # Generators in EI but not in EIA
  in_ei_only = ei_gens_copy %>% filter(is.na(eia_match_idx))
  
  # Generators in EIA but not in EI
  matched_eia_indices = in_both$eia_match_idx[!is.na(in_both$eia_match_idx)]
  in_eia_only = eia_gens %>% 
    filter(!(row_number() %in% matched_eia_indices))
  
  # Calculate summary statistics
  total_ei_capacity = sum(ei_gens$Rating, na.rm = TRUE)
  total_eia_capacity = sum(eia_gens$nameplate_capacity, na.rm = TRUE)
  matched_ei_capacity = if(nrow(in_both) > 0) sum(in_both$Rating, na.rm = TRUE) else 0
  matched_eia_capacity = if(nrow(in_both) > 0) sum(in_both$eia_capacity, na.rm = TRUE) else 0
  mean_capacity_diff = if(nrow(in_both) > 0) mean(in_both$capacity_diff, na.rm = TRUE) else 0
  
  # Print summary
  cat("\n", paste(rep("=", 60), collapse = ""), "\n")
  cat("Fleet Comparison Results\n")
  cat(paste(rep("=", 60), collapse = ""), "\n")
  cat("Generators in both EI and EIA:", nrow(in_both), "\n")
  cat("Generators in EI only:", nrow(in_ei_only), "\n")
  cat("Generators in EIA only:", nrow(in_eia_only), "\n")
  cat("\nCapacity Summary:\n")
  cat("  Total EI capacity:", round(total_ei_capacity, 2), "MW\n")
  cat("  Total EIA capacity:", round(total_eia_capacity, 2), "MW\n")
  cat("  Matched EI capacity:", round(matched_ei_capacity, 2), "MW\n")
  cat("  Matched EIA capacity:", round(matched_eia_capacity, 2), "MW\n")
  cat("  Mean capacity difference:", round(mean_capacity_diff, 2), "MW\n")
  cat(paste(rep("=", 60), collapse = ""), "\n\n")
  
  # Return results as a list
  return(list(
    in_both = in_both,
    in_ei_only = in_ei_only,
    in_eia_only = in_eia_only,
    summary = list(
      total_ei_capacity = total_ei_capacity,
      total_eia_capacity = total_eia_capacity,
      matched_ei_capacity = matched_ei_capacity,
      matched_eia_capacity = matched_eia_capacity,
      mean_capacity_diff = mean_capacity_diff,
      n_matched = nrow(in_both),
      n_ei_only = nrow(in_ei_only),
      n_eia_only = nrow(in_eia_only)
    )
  ))
}
```

## Load Data

```{r load-data, results='hide'}
# Load EI generators
VA_ei_gens = read_csv("VA_original_EI.csv")

# Filter renewable dispatch generators
ei_rd_gens = VA_ei_gens %>% 
  filter(Type == "RenewableDispatch")

# Load EIA renewable generators
eia_rd = read_csv("renewable_VA_EIA.csv")
```

## Data Overview

### EI Renewable Generators

```{r ei-overview}
cat("Total EI renewable generators:", nrow(ei_rd_gens), "\n")
cat("Total EI renewable capacity:", round(sum(ei_rd_gens$Rating), 2), "MW\n")
cat("Available generators:", sum(ei_rd_gens$Available), "\n")
cat("Unavailable generators:", sum(!ei_rd_gens$Available), "\n")
```

**Total EI renewable generators:** `r nrow(ei_rd_gens)`

**Total EI renewable capacity:** `r round(sum(ei_rd_gens$Rating), 2)` MW

**Available generators:** `r sum(ei_rd_gens$Available)`

**Unavailable generators:** `r sum(!ei_rd_gens$Available)`

### EIA Renewable Generators

```{r eia-overview}
cat("Total EIA renewable generators:", nrow(eia_rd), "\n")
cat("Total EIA renewable capacity:", round(sum(eia_rd$nameplate_capacity), 2), "MW\n")
```

## Run Comparison

We use a coordinate tolerance of 0.02 degrees (approximately 2.2 km) to match generators.

```{r run-comparison}
renewable_results = compare_generator_fleets(
  ei_rd_gens, 
  eia_rd, 
  coord_tolerance = 0.02
)
```

## Results

### Summary Statistics

```{r summary-table}
summary_df = data.frame(
  Metric = c("Generators in both datasets", 
             "Generators in EI only", 
             "Generators in EIA only",
             "Total EI capacity (MW)",
             "Total EIA capacity (MW)",
             "Matched EI capacity (MW)",
             "Matched EIA capacity (MW)",
             "Mean capacity difference (MW)"),
  Value = c(
    renewable_results$summary$n_matched,
    renewable_results$summary$n_ei_only,
    renewable_results$summary$n_eia_only,
    round(renewable_results$summary$total_ei_capacity, 2),
    round(renewable_results$summary$total_eia_capacity, 2),
    round(renewable_results$summary$matched_ei_capacity, 2),
    round(renewable_results$summary$matched_eia_capacity, 2),
    round(renewable_results$summary$mean_capacity_diff, 2)
  )
)

kable(summary_df, caption = "Renewable Generator Fleet Comparison Summary")
```

### Matched Generators

```{r matched-table}
if (nrow(renewable_results$in_both) > 0) {
  matched_display = renewable_results$in_both %>%
    dplyr::select(GenName, Rating, ei_lat, ei_lon, eia_lat, eia_lon, 
           eia_capacity, capacity_diff, Available) %>%
    head(20)
  
  kable(matched_display, 
        caption = "Sample of Matched Generators (first 20)",
        digits = 4)
}
```

### Capacity Difference Distribution

```{r capacity-plot, fig.width=10, fig.height=6}
if (nrow(renewable_results$in_both) > 0) {
  ggplot(renewable_results$in_both, aes(x = capacity_diff)) +
    geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(title = "Distribution of Capacity Differences (EI - EIA)",
         x = "Capacity Difference (MW)",
         y = "Count") +
    theme_minimal()
}
```

### Geographic Distribution

```{r geo-plot, fig.width=12, fig.height=8}
if (nrow(renewable_results$in_both) > 0) {
  ggplot() +
    geom_point(data = renewable_results$in_both, 
               aes(x = ei_lon, y = ei_lat), 
               color = "blue", alpha = 0.6, size = 3) +
    geom_point(data = renewable_results$in_both, 
               aes(x = eia_lon, y = eia_lat), 
               color = "red", alpha = 0.6, size = 3) +
    geom_segment(data = renewable_results$in_both,
                 aes(x = ei_lon, y = ei_lat, 
                     xend = eia_lon, yend = eia_lat),
                 alpha = 0.3) +
    labs(title = "Geographic Matching of Generators",
         subtitle = "Blue = EI location, Red = EIA location",
         x = "Longitude",
         y = "Latitude") +
    theme_minimal()
}
```

### EI-Only Generators

These generators exist in EI but not in EIA (potentially unavailable or retired):

```{r ei-only}
cat("Number of EI-only generators:", nrow(renewable_results$in_ei_only), "\n")
cat("Unavailable generators in EI-only:", 
    sum(!renewable_results$in_ei_only$Available), "\n")

if (nrow(renewable_results$in_ei_only) > 0) {
  ei_only_display = renewable_results$in_ei_only %>%
    dplyr::select(GenName, Rating, Lat, Lon, Available) %>%
    head(20)
  
  kable(ei_only_display, 
        caption = "Sample of EI-Only Generators (first 20)")
}
```

### EIA-Only Generators

These generators exist in EIA but not in EI:

```{r eia-only}
cat("Number of EIA-only generators:", nrow(renewable_results$in_eia_only), "\n")

if (nrow(renewable_results$in_eia_only) > 0) {
  eia_only_display = renewable_results$in_eia_only %>%
    dplyr::select(plant, technology, nameplate_capacity, lat, lon) %>%
    head(20)
  
  kable(eia_only_display, 
        caption = "Sample of EIA-Only Generators (first 20)")
}
```

## Export Results

```{r export-results, eval=FALSE}
# # Export matched generators
# write_csv(renewable_results$in_both, "matched_renewable_generators.csv")
# 
# # Export EI-only generators
# write_csv(renewable_results$in_ei_only, "ei_only_renewable_generators.csv")
# 
# # Export EIA-only generators
# write_csv(renewable_results$in_eia_only, "eia_only_renewable_generators.csv")
```

## Conclusion

This comparison identified:

- **`r renewable_results$summary$n_matched`** matched generators between EI and EIA
- **`r renewable_results$summary$n_ei_only`** generators only in EI (may be unavailable)
- **`r renewable_results$summary$n_eia_only`** generators only in EIA (not in the system model)

The mean capacity difference for matched generators is **`r round(renewable_results$summary$mean_capacity_diff, 2)` MW**.